### 2021-02-09 exploratory/exploratory.Rmd
add exploratory analysis

---
title: "OptiFit GLNE Exploratory"
author: "Courtney R Armour"
date: "February 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

library(tidyverse)
library(caret)
library(kableExtra)
library(pROC)
```

## Percent of reads mapped with OptiFit

```{r reads_mapped}

counts <- read_csv("./counts/all_counts.csv",col_names = c("sample","count"))

counts %>% 
  mutate(pct_mapped = count/10000) %>% 
  ggplot(aes(x=pct_mapped)) +
    geom_histogram(fill="grey",color="black") +
    theme_bw() + 
    scale_x_continuous(labels=scales::percent) +
    xlab("Percent of reads mapped")

# counts %>% 
#   mutate(pct_mapped = count/10000) %>% 
#   ggplot(aes(x="",y=pct_mapped)) +
#     geom_boxplot(fill="grey",color="black") +
#     theme_bw() +
#     xlab("")
```

## MCC OptiFit vs OptiClust

```{r mcc}

optifit_in_mcc <- read_tsv("./optifit_in_mcc.tsv")
optifit_out_mcc <- read_tsv("./optifit_out_mcc.tsv")
opticlust_mcc <- read_tsv("./opticlust_mcc.tsv")

mcc_data <- tibble(type=c(rep("OptiClust_489",nrow(optifit_out_mcc)),
                          rep("OptiFit_1",nrow(optifit_in_mcc)),
                          rep("OptiClust_all",nrow(opticlust_mcc))),
                   mcc=c(optifit_out_mcc$mcc,optifit_in_mcc$mcc,opticlust_mcc$mcc))

mcc_data$type <- factor(mcc_data$type,levels=c("OptiClust_all","OptiClust_489","OptiFit_1"))
mcc_data %>% 
  ggplot(aes(x=type,y=mcc,color=type)) + 
    geom_jitter(height=0) +
    xlab("") +
    theme_bw()

```

# Model performance during training
```{r data}

optifit_summary <- read_tsv("../data/learning/summary/optifit/model_results.tsv",
                            col_types = cols(sample = col_character(),
                                             cv_auc = col_double(),
                                             cancer = col_double(),
                                             normal = col_double(),
                                             dx_diff = col_double(),
                                             dx_flag = col_factor(),
                                             pred_class = col_number(),
                                             meta_class = col_number()
                                             ))  %>% 
  mutate(pred_class = case_when(pred_class == 0 ~ "normal",
                                TRUE ~ "cancer")) %>% 
  mutate(meta_class = case_when(meta_class == 0 ~ "normal",
                                TRUE ~ "cancer"))

opticlust_summary <- read_tsv("../data/learning/summary/opticlust/model_results.tsv",
                              col_types = cols(sample = col_character(),
                                               cv_auc = col_double(),
                                               cancer = col_double(),
                                               normal = col_double(),
                                               dx_diff = col_double(),
                                               dx_flag = col_factor(),
                                               pred_class = col_factor(),
                                               meta_class = col_factor()
                                               )) %>% 
  mutate(pred_class = case_when(pred_class == 0 ~ "normal",
                                TRUE ~ "cancer")) %>% 
  mutate(meta_class = case_when(meta_class == 0 ~ "normal",
                                TRUE ~ "cancer"))

opticlust_summary %>% mutate(method="OptiClust") %>% 
  bind_rows(optifit_summary %>% mutate(method="OptiFit")) %>% 
  ggplot(aes(x=method,y=cv_auc,fill=method)) +
    geom_boxplot() +
    #geom_jitter(height=0,alpha = 0.5) +
    labs(title = "Cross Validation Performance") +
    theme_bw() +
    theme(legend.position = "none") 
ggsave("./figures/cv_auc_performance.png")

```

# Prediction Probability 
```{r }

opticlust_summary %>% 
  ggplot(aes(x=cancer,fill=meta_class)) +
    geom_histogram(color="black",alpha=0.8,binwidth = 0.05) + theme_bw() +
    labs(xlab = "Probability of Cancer") + 
    facet_wrap(~meta_class) +
    xlab("probability of cancer") +
    geom_vline(xintercept = 0.5,lty="dashed") +
    scale_y_continuous(expand=c(0,0)) +
    scale_fill_manual(values=c("red","blue"),guide=NULL) +
    ggtitle("OptiClust")

optifit_summary %>% 
  ggplot(aes(x=cancer,fill=meta_class)) +
    geom_histogram(color="black",alpha=0.8,binwidth = 0.05) + theme_bw() +
    labs(xlab = "Probability of Cancer") + 
    facet_wrap(~meta_class) +
    xlab("probability of cancer") +
    geom_vline(xintercept = 0.5,lty="dashed") +
    scale_y_continuous(expand=c(0,0)) +
    scale_fill_manual(values=c("red","blue"),guide=NULL) +
    ggtitle("OptiFit")

```

### Difference in prediction probablity  
```{r hist}

# ggplot(opticlust_summary,aes(x=dx_diff)) +
#   geom_histogram(binwidth = 0.05,color="black",fill="grey") +
#   labs(title = "OptiClust: Difference in DX Prediction",
#        x="Difference in Probability") +
#   theme_bw() +
#   scale_x_continuous(expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0))
# 
# ggplot(optifit_summary,aes(x=dx_diff)) +
#   geom_histogram(binwidth = 0.05,color="black",fill="grey") +
#   labs(title = "OptiFit: Difference in DX Prediction",
#        x="Difference in Probability") +
#   theme_bw() +
#   scale_x_continuous(expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0))

opticlust_summary %>% mutate(method="OptiClust") %>% 
  bind_rows(optifit_summary %>% mutate(method="OptiFit")) %>% 
  ggplot(aes(x=dx_diff)) +
    geom_histogram(binwidth = 0.05,color="black",fill="grey") +
    labs(title = "Difference in DX Prediction",
         x="Difference in Probability") +
    theme_bw() +
    scale_x_continuous(expand=c(0,0)) +
    scale_y_continuous(expand=c(0,0)) +
    facet_wrap(~method)

```

# Summary of performance with a threshold of 0.5
```{r confusionMatrix}

opticlust_cm <- confusionMatrix(data=factor(opticlust_summary$pred_class),reference = factor(opticlust_summary$meta_class))
optifit_cm <- confusionMatrix(data=factor(optifit_summary$pred_class),reference = factor(optifit_summary$meta_class))

kable(opticlust_cm$table,caption="OptiClust Confusion Matrix") %>% 
  kable_styling(full_width = F,position = "float_left")
kable(optifit_cm$table,caption="OptiFit Confusion Matrix") %>% 
  kable_styling(full_width = F,position = "left")

cm_summary <- round(opticlust_cm$byClass,digits=3) %>% 
  bind_rows(round(optifit_cm$byClass,digits=3)) %>% 
  mutate(Method=c("OptiClust","OptiFit")) %>% 
  select(Method,everything())

kable(cm_summary,align="c",booktabs=TRUE) %>% 
  kable_classic()
```

### ROC plot

```{r roc}

#opticlust
response <- ifelse(opticlust_summary$meta_class == "cancer",1,0)
predictor <- opticlust_summary$cancer

clust_roc <- roc(response,predictor)
opticlust_thr <- coords(clust_roc, "best", ret = "threshold",transpose = TRUE)

#optifit
response <- ifelse(optifit_summary$meta_class == "normal",0,1)
predictor <- optifit_summary$normal

fit_roc <- roc(response,predictor)
optifit_thr <- coords(fit_roc, "best", ret = "threshold",transpose = TRUE)

#plot
ggroc(list(OptiClust=clust_roc,OptiFit=fit_roc),legacy.axes=TRUE) +
  theme_bw() + 
  geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1),
                 color="darkgrey", linetype="dashed") +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0))

```

OptiClust AUC: `r round(clust_roc$auc,digits=3)`  
OptiFit AUC: `r round(fit_roc$auc,digits=3)`

Optimal threshold for OptiClust: `r round(opticlust_thr,digits=3)`  
Optimal threshold for OptiFit: `r round(optifit_thr,digits=3)`