### 2022-02-21 exploratory/exploratory.Rmd
updated exploratory

---
title: "OptiFit GLNE Exploratory"
author: "Courtney R Armour"
date: "Updated: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

library(tidyverse)
library(kableExtra)
#library(caret)
#library(pROC)
```

### Table of Contents

-   [Project Summary](#summary)
-   [Overview of data splits](#splits)
-   [MCC Optifit vs OptiClust](#mcc)
-   [Hyperparameter Performance](#hp_performance)
-   [Performance during training](#cv_auc)
-   [Performance on test data](#test_auc)
-   [Average ROC curves](#avg_roc)
-   [Sensitivity at Specificity Values](#sens_xspec)
-   [Correct Classification Frequency](#class_correct_freq)
-   [Prediction Probabilities](#pred_prob)
-   [MCC vs Sensitivity/Specificity](#mcc_senspec)

### Project Summary {#summary}
...explanation of project....

### Overview of data splits {#splits}

Since the random 80% training 20% testing splits are generated 100 times, we anticipate that if the samples are randomly assigned each one should appear in the training set about 80 of the 100 times (80%) and testing about 20 of the 100 times (20%). The plot below visualizes how many times each sample appears in the training and testing datasets.

```{r plot_splits,out.width="50%"}
knitr::include_graphics("../results/figures/view_splits.png")
```

How often do samples appear in the test set together?

```{r split_together_freq}

split_count <- read_csv("../results/tables/splitTogetherFrequency.csv",col_types = cols(Group=col_character(),
                                                                                        .default = col_double()))
quant_overlap <- function(table,group1,group2){
  counts <- table %>% 
    filter(Group == c(group1,group2)) %>% 
    select(-Group) %>% 
    colSums()

  bothTest = length(which(counts == 2))
  bothTrain = length(which(counts == 0))

  tibble(group1 = group1,
         group2 = group2,
         bothTest = bothTest,
         bothTrain = bothTrain)
}

samps <- split_count %>% pull(Group)

# togetherFreq <- expand_grid(x=1:length(samps),
#             y=1:length(samps)) %>% 
#   filter(x < y) %>% 
#   mutate(group1 = samps[x],
#          group2 = samps[y]) %>% 
#   select(-x,-y) %>% 
#   group_by(group1,group2) %>% 
#   summarize(quant_overlap(split_count,group1,group2),
#             .groups="drop")
# write_csv(togetherFreq,"./data/togetherFreq.csv")
togetherFreq <- read_csv("./data/togetherFreq.csv",col_types = cols(group1=col_character(),
                                                                    group2=col_character(),
                                                                    .default=col_double()))

meanBothTest = round(mean(togetherFreq$bothTest))
meanBothTrain = round(mean(togetherFreq$bothTrain))

neverBothTest = togetherFreq %>% 
  filter(bothTest == 0) %>% 
  nrow()
neverBothTestFreq = round(neverBothTest / nrow(togetherFreq),digits=2)*100

alwaysBothTrain = togetherFreq %>% 
  filter(bothTrain == 100) %>% 
  nrow()
alwaysBothTrainFreq = round(alwaysBothTrain / nrow(togetherFreq),digits=2)*100

maxBothTest = togetherFreq %>% 
  filter(bothTest == max(bothTest))
```

For each pair of samples, I quantified how many splits both samples are in the test set together. On average, samples are both in the test set for `r meanBothTest` split. A majority of sample pairs are never in the test set together (`r neverBothTestFreq`%). At most there are `r nrow(maxBothTest)` sample pairs that are both in the test set for `r maxBothTest %>% pull(bothTest) %>% unique()` splits.

```{r together_freq_plot}
# togetherFreq %>% 
#   pivot_longer(!c(group1,group2),names_to="type",values_to = "freq") %>% 
#   ggplot(aes(x=freq,fill=type)) +
#     geom_histogram(binwidth = 1) +
#     facet_wrap(.~type,nrow=2,scales="free_y") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) + 
#     scale_x_continuous(breaks = seq(0,100,10),expand=c(0,0))
# 
# meanBothTest = round(mean(togetherFreq$bothTest))
# meanBothTrain = round(mean(togetherFreq$bothTrain))

# togetherFreq %>% 
#   pivot_longer(!c(group1,group2),names_to="type",values_to = "freq") %>% 
#   ggplot(aes(x=freq,fill=type)) +
#     geom_density() +
#     facet_wrap(.~type,nrow=2,scales = "free_y") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) + 
#     scale_x_continuous(breaks = seq(0,100,10),expand=c(0,0))

togetherFreq %>% 
  ggplot(aes(x=bothTest)) +
    geom_histogram(binwidth=1,fill="#F8766D",color="black",alpha=0.8) +
    theme_bw() +
    scale_y_continuous(expand=c(0,0)) +
    scale_x_continuous(expand=c(0,0)) +
    xlab("Number of splits where sample pair are both in the test set") +
    ylab("Number of Sample Pairs")

```

I also split the pairs by diagnosis (both cancer, both normal, or mixed) to make sure there was no difference by dx.

```{r together_freq_plot2}
sample_dx <- read_csv("../data/learning/summary/merged_predictions.csv",
                      col_types = cols(cancer=col_double(),                              
                                       normal=col_double(),
                                       .default=col_character())) %>% 
  select(Group,dx) %>% 
  distinct()

togetherFreq %>% 
  inner_join(.,sample_dx,by=c("group1" = "Group")) %>% 
  rename(group1_dx = dx) %>% 
  inner_join(.,sample_dx,by=c("group2" = "Group")) %>% 
  rename(group2_dx = dx) %>% 
  select(group1,group1_dx,group2,group2_dx,bothTest,bothTrain) %>% 
  mutate(dx=case_when(group1_dx == "normal" & group2_dx == "normal" ~ "normal",
                      group1_dx == "cancer" & group2_dx == "cancer" ~ "cancer",
                      TRUE ~ "mix")) %>% 
  ggplot(aes(x=bothTest,fill=dx)) +
    geom_histogram(binwidth=1,color="black",alpha=0.8) +
    #geom_density(color="black",alpha=0.5) +
    theme_bw() +
    scale_y_continuous(expand=c(0,0)) +
    scale_x_continuous(expand=c(0,0)) +
    xlab("Number of splits where sample pair are both in the test set") +
    ylab("Number of Sample Pairs") +
    facet_wrap(.~dx,nrow=3,scales="free_y")

# togetherFreq %>% 
#   inner_join(.,sample_dx,by=c("group1" = "Group")) %>% 
#   rename(group1_dx = dx) %>% 
#   inner_join(.,sample_dx,by=c("group2" = "Group")) %>% 
#   rename(group2_dx = dx) %>% 
#   select(group1,group1_dx,group2,group2_dx,bothTest,bothTrain) %>% 
#   mutate(dx=case_when(group1_dx == "normal" & group2_dx == "normal" ~ "normal",
#                       group1_dx == "cancer" & group2_dx == "cancer" ~ "cancer",
#                       TRUE ~ "mix")) %>% 
#   ggplot(aes(x=dx,y=bothTest,color=dx)) +
#     geom_jitter(width=0.3,height=0,alpha=0.7)
```

### Percent of reads mapped with OptiFit {#pct_mapped}

```{r pct_mapped}
# pct_mapped <- read_tsv("../results/tables/fraction_reads_mapped.tsv") %>% 
#   rename(split=sample) %>% 
#   mutate(algorithm="optifit") %>% 
#   mutate(split=paste0("split_",split))
# 
# # ggplot(pct_mapped,aes(x=algorithm,y=fraction_mapped)) + 
# #   geom_jitter(width=0.2,color="blue",size=3,alpha = 0.5) +
# #   stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
# #   theme_bw() +
# #   xlab("") + ylab("Fraction of Reads Mapped with OptiFit") +
# #   scale_y_continuous(limits=c(.98,1),labels = scales::percent_format(accuracy=0.1))
# 
# 
# highest_pct_mapped <- pct_mapped %>% 
#   filter(fraction_mapped == max(fraction_mapped)) %>% 
#   pull(split)
# lowest_pct_mapped <- pct_mapped %>% 
#   filter(fraction_mapped == min(fraction_mapped)) %>% 
#   pull(split)
# 
# ggplot(pct_mapped,aes(x=fraction_mapped)) + 
#   geom_histogram(color="black",fill="grey") +
#   theme_bw() +
#   xlab("Percent of Reads Mapped with OptiFit") +
#   scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
#   scale_y_continuous(expand=c(0,0)) +
#   geom_text(x=max(pct_mapped$fraction_mapped),y=1.5,label=highest_pct_mapped,hjust=0.25,size=3) +
#   geom_text(x=min(pct_mapped$fraction_mapped),y=1.5,label=lowest_pct_mapped,hjust=0.5,size=3)

```

### MCC OptiFit vs OptiClust {#mcc}

```{r mcc}
merged_mcc <- read_csv("../data/learning/summary/merged_MCC.csv") %>% 
  mutate(label = case_when(algorithm == "opticlust" ~ algorithm,
                           TRUE ~ paste0(algorithm,"_",subset))) %>% 
  mutate(Group=case_when(algorithm == "opticlust" ~ "OptiClust",
                         subset == "train" ~ "OptiFit Reference",
                         state == "fit" ~ "Optifit Fit without Reference",
                         TRUE ~ "Optifit Fit with Reference")) %>% 
  mutate(label = factor(label,levels=c("opticlust","optifit_train","optifit_test"),
                        labels = c("OptiClust","OptiFit Train","OptiFit Test")))

merged_mcc %>% 
  ggplot(aes(x=label,y=mcc,color=Group)) +
    geom_jitter(height=0,width=0.2,alpha=0.5,size=3) +
    theme_bw() +
    xlab("") + ylab("MCC") +
    #scale_y_continuous(expand=c(0,0),limits=c(0.85,0.91)) +
    theme(panel.grid.major.y = element_line(color="grey85"),
          panel.grid.minor.y = element_line(color="grey85"),
          panel.grid.major.x = element_blank()) +
    geom_text(aes(label,mcc,label=split),data=subset(merged_mcc, mcc < 0.7),color="black",hjust=0) #+
    # geom_text(aes(label,mcc,label=split),data=subset(merged_mcc, split == lowest_pct_mapped),color="black",hjust=0) +
    # geom_text(aes(label,mcc,label=split),data=subset(merged_mcc, split == highest_pct_mapped),color="black",hjust=0)

```
```{r opticlust_20_mcc}

opticlust_20_mcc <- read_csv("../results/tables/opticlust_20_mcc.csv") %>% 
  mutate(Group="OptiClust 20%")

merged_mcc %>% 
  filter(algorithm=="optifit",subset=="test") %>% 
  select(mcc,split,algorithm,Group) %>% 
  bind_rows(opticlust_20_mcc) %>% 
  ggplot(aes(x=algorithm,y=mcc,color=Group)) +
    geom_jitter(height=0,width=0.2,alpha=0.5,size=3) +
    theme_bw() +
    xlab("") + ylab("MCC") +
    theme(panel.grid.major.y = element_line(color="grey85"),
          panel.grid.minor.y = element_line(color="grey85"),
          panel.grid.major.x = element_blank())

```

```{r pct_map_v_mcc_ref}
### Percent Mapped VS MCC (without reference) {#pct_v_mcc_ref}
# 
# sub_pct <- pct_mapped %>% 
#   select(split,fraction_mapped)
# 
# mcc_pct <- merged_mcc %>% 
#   filter(algorithm == "optifit" & subset == "test",state == "fit") %>% 
#   select(split,mcc) %>% 
#   inner_join(., sub_pct ,by="split") 
# 
# mcc_pct %>% 
#   ggplot(aes(x=fraction_mapped,y=mcc)) +
#     geom_point(size=2,alpha = 0.5) +
#     theme_bw() +
#     xlab("Percent of Reads Mapped with OptiFit") +
#     ylab("OptiFit MCC (printref=f)") +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (fraction_mapped > 0.997 | fraction_mapped < 0.98)),
#               color="black",vjust=-1,size=3) +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (mcc > 0.775 | mcc < 0.7)),
#               color="black",vjust=-1,size=3)
```

```{r pct_map_v_mcc}
### Percent Mapped VS MCC (WITH reference) {#pct_v_mcc_noref}

# sub_pct <- pct_mapped %>% 
#   select(split,fraction_mapped)
# 
# mcc_pct <- merged_mcc %>% 
#   filter(algorithm == "optifit" & subset == "test",state == "combo") %>% 
#   select(split,mcc) %>% 
#   inner_join(., sub_pct ,by="split") 
# 
# mcc_pct %>% 
#   ggplot(aes(x=fraction_mapped,y=mcc)) +
#     geom_point(size=2,alpha = 0.5) +
#     theme_bw() +
#     xlab("Percent of Reads Mapped with OptiFit") +
#     ylab("OptiFit MCC (printref=t)") +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (fraction_mapped > 0.997 | fraction_mapped < 0.98)),
#               color="black",vjust=-1,size=3) +
#     geom_text(aes(fraction_mapped,mcc,label=split),
#               data=subset(mcc_pct, (mcc > 0.8961 | mcc < 0.8905)),
#               color="black",vjust=-1,size=3)
```

### HP performance {#hp_performance}

```{r hp}
# hp_performance <- read_csv("../data/learning/summary/merged_HP.csv",
#                            col_types = cols(split=col_character(),
#                                             algorithm=col_character(),
#                                             .default=col_double()))
# hp_performance %>% 
#   group_by(mtry,algorithm) %>% 
#   summarise("mean_AUC" = mean(AUC),
#             "sd_AUC" = sd(AUC),
#             ymin_metric = mean_AUC - sd_AUC,
#             ymax_metric = mean_AUC + sd_AUC)
# ggplot(hp_performance,aes(x=mtry,y=AUC,color=algorithm)) +
#   geom_point() +
#   facet_wrap(.~algorithm)
knitr::include_graphics("../results/figures/hp_performance.png")
```

### Model performance during training {#cv_auc}

```{r cv_auc}
#allCV <- read_csv("../data/learning/summary/merged_CV.csv")
allCV <- read.csv("../data/learning/summary/merged_performance.csv")

meds <- allCV %>% 
  group_by(algorithm) %>% 
  summarise(median_CV_AUC = round(median(cv_metric_AUC),digits=3))

ggplot(allCV,aes(x=algorithm,y=cv_metric_AUC,color=algorithm)) +
  geom_jitter(width = 0.2,size=3,alpha = 0.5) +
  stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  theme_bw() +
  ylab("Cross-Validation AUC") +
  geom_text(data=meds,aes(x=algorithm,y=median_CV_AUC,label=median_CV_AUC),
            color="black",hjust=-2)

allCV %>% 
  select(cv_metric_AUC,split,algorithm) %>% 
  pivot_wider(id_cols = split,names_from = algorithm,values_from = cv_metric_AUC) %>% 
  mutate(diff = opticlust-optifit) %>% 
  mutate(change_from_opticlust = case_when(diff >= 0.01 ~ "decrease",
                                           diff <= -0.01 ~ "increase",
                                           TRUE ~ "less than |0.01|")) %>% 
  pivot_longer(cols = c("opticlust","optifit"),names_to="algorithm",values_to = "cv_metric_AUC") %>% 
  ggplot(aes(x=algorithm,y=cv_metric_AUC,color=change_from_opticlust,group=split)) +
    geom_point(size=3,alpha = 0.5) +
    geom_line(alpha = 0.5) +
    theme_bw() +
    scale_color_manual(values = c("red","blue","grey")) +
    ylab("Cross-Validation AUC")

# ggplot(allCV,aes(x=algorithm,y=cv_metric_AUC,color=split,group=split)) +
#     geom_point(size=3,alpha = 0.5) +
#     geom_line() +
#     theme_bw() +
#     ylab("Cross-Validation AUC") +
#     theme(legend.position = "none")
```

### Model performance on test data {#test_auc}

```{r sens_spec}
all_sens_spec <- read_csv("../data/learning/summary/all_sens_spec.csv",
                          col_types = cols(algorithm=col_character(),
                                           .default=col_double()))
```

```{r test_auc}
allAUCs <- read_csv("../data/learning/summary/merged_performance.csv",
                    col_types = cols(algorithm=col_character(),
                                     method=col_character(),
                                     split=col_character(),
                                     .default=col_double())) 
meds <- allAUCs %>% 
  group_by(algorithm) %>% 
  summarise(median_AUC = round(median(AUC),digits=3))

ggplot(allAUCs,aes(x=algorithm,y=AUC,color=algorithm)) +
  geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) + 
  stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  theme_bw() +
  ylab("AUC") +
  xlab("") +
  theme(legend.position = "top") +
  geom_text(data=meds,aes(x=algorithm,y=median_AUC,label=median_AUC),
            color="black",hjust=-2.8)
```

```{r allAUC}

allAUCs <- read_csv("../data/learning/summary/merged_performance.csv",
                    col_types = cols(algorithm=col_character(),
                                     method=col_character(),
                                     split=col_character(),
                                     .default=col_double())) 
meds <- allAUCs %>% 
  select(cv_metric_AUC,AUC,algorithm) %>% 
  rename(CV=cv_metric_AUC,
         Test=AUC) %>% 
  pivot_longer(!algorithm,names_to="type",values_to = "AUC") %>% 
  group_by(algorithm,type) %>% 
  summarise(median_AUC = round(median(AUC),digits=3))

allAUCs %>% 
  select(cv_metric_AUC,AUC,algorithm) %>% 
  rename(CV=cv_metric_AUC,
         Test=AUC) %>% 
  pivot_longer(!algorithm,names_to="type",values_to = "AUC") %>% 
  ggplot(aes(x=type,y=AUC,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.6,size=2) + 
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    facet_grid(.~algorithm) +
    theme(legend.position = "top") +
    geom_text(data=meds,aes(x=type,y=median_AUC,label=median_AUC),
              color="black",hjust=-1.2)

```

### Averaged ROC Curve {#avg_roc}

```{r avg_roc}
# calculate average/sd sensitivity
avg_sens <- all_sens_spec %>%
  group_by(specificity,algorithm) %>%
  summarise(mean_sensitivity = mean(sensitivity),
            sd_sensitivity = sd(sensitivity),
            .groups = "drop") %>%
  mutate(upper_sens = mean_sensitivity + sd_sensitivity,
         lower_sens = mean_sensitivity - sd_sensitivity) %>%
  mutate(upper_sens = case_when(upper_sens > 1 ~ 1,
                                TRUE ~ upper_sens),
         lower_sens = case_when(upper_sens < 0 ~ 0,
                                TRUE ~ lower_sens)) %>%
  mutate(fpr = 1-specificity)

#calculate mean/median
mean_med <- allAUCs %>% 
  group_by(algorithm) %>% 
  summarize(median_AUC = round(median(AUC),3),
            mean_AUC = round(mean(AUC),3))

of_mean <- mean_med %>% filter(algorithm=="optifit") %>% pull(mean_AUC)
oc_mean <- mean_med %>% filter(algorithm=="opticlust") %>% pull(mean_AUC)

# plot all levels together
avg_sens %>%
  ggplot(aes(x=fpr,y=mean_sensitivity,
             ymin=lower_sens,ymax=upper_sens)) +
  geom_line(aes(color=algorithm),alpha=0.7,lwd=1) +
  geom_ribbon(aes(fill=algorithm),alpha=0.3) +
  coord_equal() +
  geom_abline(intercept = 0,lty="dashed",color="grey50") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  xlab("False Positive Rate") +
  ylab("Average True Positive Rate") +
  #facet_wrap(~level,ncol=3) +
  #scale_x_reverse(expand = c(0,0),labels = scales::percent_format(accuracy = 1)) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text = element_text(size=11),
        axis.title = element_text(size=12),
        axis.text.x = element_text(vjust = -0.5),
        axis.title.x = element_text(vjust= -0.5)) +
  guides(color = guide_legend(nrow = 1,title=""),
         fill = guide_legend(nrow = 1,title="")) +
  scale_color_manual(values=c("brown3","deepskyblue3")) +
  geom_text(x=0.7,y=0.2,label=paste0("OptiClust Mean AUC: ",oc_mean),size=3.5) +
  geom_text(x=0.7,y=0.15,label=paste0("OptiFit Mean AUC: ",of_mean),size=3.5)

```

### Sensitivity at Specificity Thresholds {#sens_xspec}
```{r sens_xspec}

fun_median <- function(x){
  return(data.frame(y=round(median(x),digits=3),label=round(median(x,na.rm=T),digits=3)))
}

spec = 0.9
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)

spec = 0.5
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)

spec = 0.1
all_sens_spec %>% 
  filter(specificity == spec) %>% 
  ggplot(aes(x=algorithm,y=sensitivity,color=algorithm)) +
    geom_jitter(height=0,width=0.2,alpha = 0.7,size=2) +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    theme_bw() +
    ylab("AUC") +
    xlab("") +
    theme(legend.position = "top") +
    ggtitle(paste0("Sensitivity at ",spec*100,"% Specificity")) +
    stat_summary(fun.data=fun_median, geom="text",show.legend = FALSE,color="black",hjust=-2.2)
```

### Correct Classification Frequency {#class_correct_freq}

For each split, a cutoff of 0.5 was used to determine whether each sample was correctly or incorrectly classified. This data was merged for all splits to quantify the percent of splits each sample was correctly classified in. For normal samples it is the frequency that sample was a true negative and for cancer samples its the frequency that sample was a true positive. For example, if a cancer sample had a probabily of being cancer greater than 0.5 for 10 of the 20 splits it was in the test set for, it would have a percent correct value of 50%. I'm trying to figure out how to visualize this data...

```{r pct_correct_hist, out.width="50%"}
pct_correct <- read_csv("../results/tables/pct_class_correct.csv")

# pct_correct %>%
#   ggplot(aes(x=factor(algorithm,levels=c("opticlust","optifit"),labels=c("OptiClust","OptiFit")),
#              y=pct_correct,group=Group,color=dx)) +
#     geom_jitter(height=0,width=0.2,alpha=0.7,size=1.8) +
#     theme_bw() + xlab("") +
#     ylab("Percent of splits sample is correctly classified (TP or TN)") +
#     scale_y_continuous(labels = scales::percent_format(accuracy=1),
#                        breaks = seq(0,1,0.1)) +
#     theme(panel.grid.major.y = element_line(color="grey85")) +
#     scale_color_manual(values=c("brown3","deepskyblue3"),name="Actual Diagnosis")
#

pct_correct %>%
  ggplot(aes(x=pct_correct,fill=algorithm)) +
  geom_histogram(binwidth=0.1,color="black",alpha=0.5,position = "identity") +
  facet_grid(.~dx) +
  theme_bw() +
  xlab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(legend.position = "top")

pct_correct %>%
  ggplot(aes(x=pct_correct,fill=algorithm)) +
  geom_histogram(binwidth=0.1,color="black",alpha=0.5,position = "dodge") +
  facet_grid(.~dx) +
  theme_bw() +
  xlab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(legend.position = "top")

pct_correct %>%
  ggplot(aes(x=pct_correct,fill=algorithm)) +
  #geom_histogram(binwidth=0.1,color="black",alpha=0.8) +
  geom_density(alpha=0.5) +
  facet_grid(.~dx) +
  theme_bw() +
  xlab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(legend.position = "top")

pct_correct %>%
  ggplot(aes(x=pct_correct,color=algorithm)) +
  #geom_histogram(binwidth=0.1,color="black",alpha=0.8) +
  geom_density(alpha=0.5) +
  facet_grid(.~dx) +
  theme_bw() +
  xlab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(legend.position = "top")

```

The scatter plot below has the samples along the x-axis without labels because they are not readable. The y-axis is the percent correct. The color of the points is the algorithm. There is a separate plot for cancer samples and normal samples.

```{r pct_correct_point,fig.width=11}
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference) %>% 
  pull(Group)

pct_correct %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample") +
  theme(legend.position = "top")
```

To eliminate some noise, I removed any samples where there was no difference in the percent correct between the two algorithms.

```{r mod_pct_correct_point,fig.width=9}
#get rid of samples that are perfectly classified with both algorithms
#and samples that have the samp pct_correct
#order samples by difference in pct_correct between algorithms
samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  #filter(optifit != 1 & opticlust != 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference > 0) %>% 
  arrange(difference,optifit) %>% 
  pull(Group)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample") +
  theme(legend.position = "top")

```

### Prediction Probabilities {#pred_prob}
I selected the samples where the difference in the correct classification frequency between optifit and opticlust was greater than 0.5 as shown below.

```{r predprob_subset}

samp_order <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  #filter(optifit != 1 & opticlust != 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference > 0.5) %>% 
  arrange(difference,optifit) %>% 
  pull(Group)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")

```

For each of the `r length(samp_order)` samples in this set, I plotted the probability of cancer from each split this sample was in the test set for. The dashed line is the 0.5 probability line. Using the 0.5 threshold, a point above the line means the sample would be classified as cancer in that split, below the line normal. For the cancer samples, a point above the line is a true positive while for the normal samples a point above the line is a false positive

```{r predprob, fig.height = 10, fig.width = 9}
mergedProb <- read_csv("../data/learning/summary/merged_predictions.csv",
                       col_types = cols(cancer=col_double(),
                                        normal=col_double(),
                                        .default=col_character())) %>%
  mutate(GroupDX = paste0(Group," ",dx))

ids = as.character(samp_order)

## NORMAL
mergedProb %>%
  filter(Group %in% ids) %>%
  filter(dx == "normal") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")
## CANCER
mergedProb %>%
  filter(Group %in% ids) %>%
  mutate(Group=factor(Group,levels=ids)) %>% 
  filter(dx == "cancer") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="brown3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")

```

For comparison, here are a subset of samples where there is not a difference in the percent correctly classified between optifit and opticlust

```{r predprob_subset2}

samp_order1 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit == 1 & opticlust == 1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference,optifit) %>% 
  sample_n(10) %>% #randomly select 5
  pull(Group)

samp_order2 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit < 0.9 & opticlust < 0.9) %>% 
  filter(optifit > 0.1 & opticlust > 0.1) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  filter(difference == 0) %>% 
  sample_n(min(10,nrow(.))) %>% #randomly select 5
  arrange(optifit) %>% 
  pull(Group)

samp_order3 <- pct_correct %>% 
  pivot_wider(id_cols = c("Group","dx"),
              names_from = algorithm,
              values_from = pct_correct) %>% 
  filter(optifit == 0 & opticlust == 0) %>% 
  mutate(difference = abs(opticlust - optifit)) %>% 
  arrange(difference,optifit) %>% 
  sample_n(10) %>% #randomly select 5
  pull(Group)

samp_order <- c(samp_order1,samp_order2,samp_order3)

pct_correct %>% 
  filter(Group %in% samp_order) %>% 
  ggplot(aes(x=factor(Group,levels = samp_order),y=pct_correct,color=algorithm)) +
  geom_point(alpha=0.7) +
  facet_wrap(.~dx,nrow=2,scales="free_x") +
  theme_bw() + 
  ylab("Percent of splits sample is correctly classified (TP or TN)") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Sample")

```

```{r predprob2, fig.height = 10, fig.width = 9}
mergedProb <- read_csv("../data/learning/summary/merged_predictions.csv",
                       col_types = cols(cancer=col_double(),
                                        normal=col_double(),
                                        .default=col_character())) %>%
  mutate(GroupDX = paste0(Group," ",dx))

ids = as.character(samp_order)

## NORMAL
mergedProb %>%
  filter(Group %in% ids) %>%
  filter(dx == "normal") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")
## CANCER
mergedProb %>%
  filter(Group %in% ids) %>%
  mutate(Group=factor(Group,levels=ids)) %>% 
  filter(dx == "cancer") %>% 
  ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
    geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
    facet_wrap(.~GroupDX) +
    theme_bw() +
    geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
    ylab("Probability of SRN") + xlab("") +
    scale_y_continuous(limits=c(0,1),breaks=seq(0.2,0.8,0.2)) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          strip.background = element_rect(fill="brown3", size=1, color="black"),
          strip.text.x = element_text(face="bold", color="white"),
          legend.position = "top")

```

## Does mcc correlate with sensitivity/specificity? {#mcc_senspec}

OptiFit MCC values from just the fit data compared to sensitivity/specificity (threshold = 0.5)

```{r mcc_sensspec}
library(caret)
library(assertthat)

get_sensspec_thrsh <- function(senspec,threshold){
  assert_that(is.numeric(threshold))

  senspec %>%
    mutate(prediction = case_when(cancer >= threshold ~ "cancer",
                                  TRUE ~ "normal")) %>%
    select(-cancer,-normal) %>% 
    group_by(split,algorithm) %>% 
    summarize()

  senspec %>%
    mutate(prediction = case_when(cancer >= 0.5 ~ "cancer",
                                  TRUE ~ "normal")) %>%
    mutate(dx = factor(dx,levels=c("normal","cancer")),
           prediction = factor(prediction,levels=c("normal","cancer"))) %>% 
    select(-cancer,-normal) %>% 
    mutate(category = case_when(dx == "normal" & prediction == "normal" ~ "TN",
                                dx == "cancer" & prediction == "cancer" ~ "TP",
                                dx == "cancer" & prediction == "normal" ~ "FN",
                                dx == "normal" & prediction == "cancer" ~ "FP")) %>% 
    group_by(split,algorithm) %>% 
    count(category) %>% 
    pivot_wider(id_cols=c("split","algorithm"),names_from = category,values_from = n) %>% 
    group_by(split,algorithm) %>% 
    summarise(sensitivity = TP / (TP+FN),
              specificity = TN / (TN+FP))

}

senspec_50 <- get_sensspec_thrsh(read_csv("../data/learning/summary/merged_predictions.csv"),0.5)

# senspec_50 %>% 
#   filter(algorithm == "optifit")

merged_mcc %>% 
  filter(label=="OptiFit Test" & state=="fit") %>% 
  select(-subset,-Group) %>% 
  inner_join(senspec_50 %>% filter(algorithm == "optifit"),by=c("split","algorithm")) %>% 
  ggplot(aes(x=mcc,y=sensitivity)) +
    geom_point() +
    theme_bw()

merged_mcc %>% 
  filter(label=="OptiFit Test" & state=="fit") %>% 
  select(-subset,-Group) %>% 
  inner_join(senspec_50 %>% filter(algorithm == "optifit"),by=c("split","algorithm")) %>% 
  ggplot(aes(x=mcc,y=specificity)) +
    geom_point() +
    theme_bw()

```

```{r test}
# 
# get_categories <- function(filename,threshold){
#   read_csv(filename,col_types = cols(Group = col_character(),
#                                     dx = col_character(),
#                                     .default = col_double())) %>% 
#     mutate(algorithm = str_replace(filename,
#                                    "../data/learning/results/(.*)/prediction_results_split_(\\d*).csv", "\\1"),
#            split = str_replace(filename,
#                                "../data/learning/results/(.*)/prediction_results_split_(\\d*).csv", "\\2"),
#            pred_dx = case_when(cancer > threshold ~ "cancer",
#                                TRUE ~ "normal")) %>% 
#     filter(Group %in% c("2005650","2009650","2025653","2671660")) %>% 
#     select(Group,dx,pred_dx,algorithm,split) %>% 
#     mutate(category = case_when(dx == "normal" & pred_dx == "normal" ~ "TN",
#                                 dx == "normal" & pred_dx == "cancer" ~ "FP",
#                                 dx == "cancer" & pred_dx == "cancer" ~ "TP",
#                                 dx == "cancer" & pred_dx == "normal" ~ "FN",
#                                 TRUE ~ "NA"))
# }
# 
# get_pct_correct <- function(algorithm, threshold){
#   #get_pred_file_list(algorithm) %>% 
#   filelist <- c("../data/learning/results/optifit/prediction_results_split_1.csv",
#                 "../data/learning/results/optifit/prediction_results_split_2.csv",
#                 "../data/learning/results/optifit/prediction_results_split_3.csv",
#                 "../data/learning/results/optifit/prediction_results_split_4.csv",
#                 "../data/learning/results/optifit/prediction_results_split_5.csv")
  # filelist <- c("../data/learning/results/optifit/prediction_results_split_1.csv",
  #               "../data/learning/results/optifit/prediction_results_split_2.csv",
  #               "../data/learning/results/optifit/prediction_results_split_3.csv",
  #               "../data/learning/results/optifit/prediction_results_split_4.csv",
  #               "../data/learning/results/optifit/prediction_results_split_5.csv")
#   filelist %>% 
#     map_dfr(get_categories,threshold) %>% 
#     select(Group,category,dx) %>%
#     group_by(Group,dx) %>% 
#     count(category) %>% 
#     pivot_wider(id_cols=c("Group","dx"),names_from=category,values_from=n,values_fill=0) %>% 
#     mutate(correct = TN + TP,
#            incorrect = FN + FP,
#            total = TN + TP + FN + FP) %>% 
#     mutate(pct_correct = correct/total) %>% 
#     select(Group,dx,pct_correct) %>% 
#     mutate(algorithm = algorithm)
# }
# 
# pct_correct <- map_dfr(algorithms, get_pct_correct, threshold=0.5) %>% 
#   mutate(Group=as.character(Group))
# 
# pct_correct %>% 
#   ggplot(aes(x=algorithm,y=pct_correct,group=Group,color=dx)) +
#     geom_jitter(height=0,width=0.2) +
#     theme_bw() +
#     ylab("Percent of splits sample is correctly classified (TP or TN)")

```

```{r }

### Prediction Probability 
# opticlust_summary %>% 
#   ggplot(aes(x=cancer,fill=meta_class)) +
#     geom_histogram(color="black",alpha=0.8,binwidth = 0.05) + theme_bw() +
#     labs(xlab = "Probability of Cancer") + 
#     facet_wrap(~meta_class) +
#     xlab("probability of cancer") +
#     geom_vline(xintercept = 0.5,lty="dashed") +
#     scale_y_continuous(expand=c(0,0)) +
#     scale_fill_manual(values=c("red","blue"),guide=NULL) +
#     ggtitle("OptiClust")

# optifit_summary %>% 
#   ggplot(aes(x=cancer,fill=meta_class)) +
#     geom_histogram(color="black",alpha=0.8,binwidth = 0.05) + theme_bw() +
#     labs(xlab = "Probability of Cancer") + 
#     facet_wrap(~meta_class) +
#     xlab("probability of cancer") +
#     geom_vline(xintercept = 0.5,lty="dashed") +
#     scale_y_continuous(expand=c(0,0)) +
#     scale_fill_manual(values=c("red","blue"),guide=NULL) +
#     ggtitle("OptiFit")

```

```{r }
### Difference in prediction probablity  
# ggplot(opticlust_summary,aes(x=dx_diff)) +
#   geom_histogram(binwidth = 0.05,color="black",fill="grey") +
#   labs(title = "OptiClust: Difference in DX Prediction",
#        x="Difference in Probability") +
#   theme_bw() +
#   scale_x_continuous(expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0))
# 
# ggplot(optifit_summary,aes(x=dx_diff)) +
#   geom_histogram(binwidth = 0.05,color="black",fill="grey") +
#   labs(title = "OptiFit: Difference in DX Prediction",
#        x="Difference in Probability") +
#   theme_bw() +
#   scale_x_continuous(expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0))

# opticlust_summary %>% mutate(method="OptiClust") %>% 
#   bind_rows(optifit_summary %>% mutate(method="OptiFit")) %>% 
#   ggplot(aes(x=dx_diff)) +
#     geom_histogram(binwidth = 0.05,color="black",fill="grey") +
#     labs(title = "Difference in DX Prediction",
#          x="Difference in Probability") +
#     theme_bw() +
#     scale_x_continuous(expand=c(0,0)) +
#     scale_y_continuous(expand=c(0,0)) +
#     facet_wrap(~method)

```

```{r }
### Summary of performance with a threshold of 0.5
# opticlust_cm <- confusionMatrix(data=factor(opticlust_summary$pred_class),reference = factor(opticlust_summary$meta_class))
# optifit_cm <- confusionMatrix(data=factor(optifit_summary$pred_class),reference = factor(optifit_summary$meta_class))

# kable(opticlust_cm$table,caption="OptiClust Confusion Matrix") %>% 
#   kable_styling(full_width = F,position = "float_left")
# kable(optifit_cm$table,caption="OptiFit Confusion Matrix") %>% 
#   kable_styling(full_width = F,position = "left")

# cm_summary <- round(opticlust_cm$byClass,digits=3) %>% 
#   bind_rows(round(optifit_cm$byClass,digits=3)) %>% 
#   mutate(Method=c("OptiClust","OptiFit")) %>% 
#   select(Method,everything())

# kable(cm_summary,align="c",booktabs=TRUE) %>% 
#   kable_classic()
```

```{r }
### ROC plot  
#opticlust
# response <- ifelse(opticlust_summary$meta_class == "cancer",1,0)
# predictor <- opticlust_summary$cancer

# clust_roc <- roc(response,predictor)
# opticlust_thr <- coords(clust_roc, "best", ret = "threshold",transpose = TRUE)

# #optifit
# response <- ifelse(optifit_summary$meta_class == "normal",0,1)
# predictor <- optifit_summary$normal

# fit_roc <- roc(response,predictor)
# optifit_thr <- coords(fit_roc, "best", ret = "threshold",transpose = TRUE)

# #plot
# ggroc(list(OptiClust=clust_roc,OptiFit=fit_roc),legacy.axes=TRUE) +
#   theme_bw() + 
#   geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1),
#                  color="darkgrey", linetype="dashed") +
#   scale_y_continuous(expand=c(0,0)) +
#   scale_x_continuous(expand=c(0,0))

```

```{r}
# OptiClust AUC: `r round(clust_roc$auc,digits=3)`  
# OptiFit AUC: `r round(fit_roc$auc,digits=3)`

# Optimal threshold for OptiClust: `r round(opticlust_thr,digits=3)`  
# Optimal threshold for OptiFit: `r round(optifit_thr,digits=3)`  
```

```{r pred_prob,fig.width=7,fig.height=7}
# library(ggforce)
# 
# mergedProb <- read_csv("../data/learning/summary/merged_predictions.csv",
#                        col_types = cols(cancer=col_double(),
#                                         normal=col_double(),
#                                         .default=col_character())) %>% 
#   mutate(GroupDX = paste0(Group," ",dx))
# 
# ids = c((mergedProb %>% filter(dx == "normal") %>% pull(Group))[1:5],
#         (mergedProb %>% filter(dx == "cancer") %>% pull(Group))[1:5])
# 
# ### NORMAL
# p <- mergedProb %>%
#   #filter(Group %in% ids) %>%
#   filter(dx == "normal") %>%
#   ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#     geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
#     facet_wrap_paginate(~GroupDX,ncol=5,nrow=5) +
#     theme_bw() +
#     geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#     ylab("Probability of SRN") + xlab("") +
#     #scale_y_continuous(limits=c(0,1),breaks=seq(0,1,0.2)) +
#     theme(axis.text.x = element_blank(),
#           axis.ticks.x = element_blank(),
#           strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"),
#           legend.position = "top")
# 
# ###
#   # plot <- mergedProb %>%
#   # filter(dx == "normal") %>%
#   # ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#   #   geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
#   #   facet_wrap_paginate(~GroupDX,ncol=5,nrow=5,page=1) +
#   #   theme_bw() +
#   #   geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#   #   ylab("Probability of SRN") + xlab("") +
#   #   #scale_y_continuous(limits=c(0,1),breaks=seq(0,1,0.2)) +
#   #   theme(axis.text.x = element_blank(),
#   #         axis.ticks.x = element_blank(),
#   #         strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
#   #         strip.text.x = element_text(face="bold", color="white"),
#   #         legend.position = "top")
#   # print(plot)
# ###
# 
# for( i in 1:n_pages(p)){
#   plot <- mergedProb %>%
#   #filter(Group %in% ids) %>%
#   filter(dx == "normal") %>%
#   ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#     geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
#     facet_wrap_paginate(~GroupDX,ncol=5,nrow=5,page=i) +
#     theme_bw() +
#     geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#     ylab("Probability of SRN") + xlab("") +
#     #scale_y_continuous(limits=c(0,1),breaks=seq(0,1,0.2)) +
#     theme(axis.text.x = element_blank(),
#           axis.ticks.x = element_blank(),
#           strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"),
#           legend.position = "top")
#   print(plot)
# }
# 
# ### CANCER
# p <- mergedProb %>%
#   #filter(Group %in% ids) %>%
#   filter(dx == "cancer") %>%
#   ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#     geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
#     facet_wrap_paginate(~GroupDX,ncol=6,nrow=5) +
#     theme_bw() +
#     geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#     ylab("Probability of SRN") + xlab("") +
#     #scale_y_continuous(limits=c(0,1),breaks=seq(0,1,0.2)) +
#     theme(axis.text.x = element_blank(),
#           axis.ticks.x = element_blank(),
#           strip.background = element_rect(fill="brown3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"),
#           legend.position = "top")
# 
# for( i in 1:n_pages(p)){
#   plot <- mergedProb %>%
#   #filter(Group %in% ids) %>%
#   filter(dx == "cancer") %>%
#   ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#     geom_jitter(width = 0.2,height=0,alpha=0.5,size=1.8) +
#     facet_wrap_paginate(~GroupDX,ncol=6,nrow=5,page=i) +
#     theme_bw() +
#     geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#     ylab("Probability of SRN") + xlab("") +
#     #scale_y_continuous(limits=c(0,1),breaks=seq(0,1,0.2)) +
#     theme(axis.text.x = element_blank(),
#           axis.ticks.x = element_blank(),
#           strip.background = element_rect(fill="brown3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"),
#           legend.position = "top")
#   print(plot)
# }
# 
#

# mergedProb %>% 
#   #filter(Group %in% ids) %>% 
#   filter(dx == "cancer") %>% 
#   ggplot(aes(x=algorithm,y=cancer,color=algorithm)) +
#     geom_jitter(width = 0.2,height=0,alpha=0.7,size=1.8) +
#     facet_wrap(~GroupDX,ncol=5) +
#     theme_bw() +
#     geom_hline(yintercept = 0.5,lty="dashed",color="grey") +
#     ylab("Probability of SRN") +
#     theme(axis.text.x = element_text(angle=45,hjust=1),
#           strip.background = element_rect(fill="brown3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"))
# 
```

```{r prob_hist,include=F}
# mergedProb %>% 
#   #filter(Group %in% ids) %>% 
#   filter(dx == "normal") %>% 
#   mutate(GroupDX = paste0(Group," ",dx)) %>% 
#   ggplot(aes(x=cancer,fill=algorithm)) +
#     geom_histogram(binwidth = 0.05,color="black") +
#     geom_vline(xintercept=0.5,lty="dashed",color="grey") +
#     facet_wrap(~GroupDX,ncol=5) +
#     xlab("Probability of SRN") +
#     ylab("Number of Splits") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) +
#     scale_x_continuous(limits=c(0,1),labels = scales::percent_format(accuracy=1)) +
#     theme(axis.text.x = element_text(angle=45),
#           strip.background = element_rect(fill="deepskyblue3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"))
# mergedProb %>% 
#   #filter(Group %in% ids) %>% 
#   filter(dx == "cancer") %>% 
#   mutate(GroupDX = paste0(Group," ",dx)) %>% 
#   ggplot(aes(x=cancer,fill=algorithm)) +
#     geom_histogram(binwidth = 0.05,color="black") +
#     geom_vline(xintercept=0.5,lty="dashed",color="grey") +
#     facet_wrap(~GroupDX,ncol=5) +
#     xlab("Probability of SRN") +
#     ylab("Number of Splits") +
#     theme_bw() +
#     scale_y_continuous(expand=c(0,0)) +
#     scale_x_continuous(limits=c(0,1),labels = scales::percent_format(accuracy=1)) +
#     theme(axis.text.x = element_text(angle=45),
#           strip.background = element_rect(fill="brown3", size=1, color="black"),
#           strip.text.x = element_text(face="bold", color="white"))

```

```{r prob_diff,include=F}

# mergedProb %>% 
#   #group_by(Group,dx,algorithm,split) %>%   
#   #filter(Group %in% ids) %>% 
#   pivot_wider(id_cols=c("Group","dx","split"),names_from = algorithm,values_from = cancer) %>% 
#   mutate(diff = opticlust - optifit) %>% 
#   ggplot(aes(x=diff,fill=dx,color=dx)) +
#     geom_histogram(alpha=0.7) +
#     facet_wrap(~Group) +
#     geom_vline(xintercept = 0,lty="dashed",color="grey") +
#     xlab("Cancer Probability Difference (OptiClust - OptiFit)") +
#     scale_y_continuous(expand=c(0,0)) +
#     theme_bw() +
#     scale_fill_manual(values=c("brown3","deepskyblue3"),name="Actual Diagnosis") +
#     scale_color_manual(values=c("brown3","deepskyblue3"),name="Actual Diagnosis")

```