### 2022-06-20 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Reference-based OTU clustering for machine learning classification},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\usepackage{multirow}
\usepackage[none]{hyphenat}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{\textbf{Reference-based OTU clustering for machine learning
classification}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{5mm}

Running title: Reference-based OTU clustering for ML classification

\vspace{10mm}

Courtney R. Armour\({^1}\), Kelly L. Sovacool\({^2}\), William L.
Close\(^{1,*}\), Begüm D. Topçuoğlu\(^{1,\#}\), Jenna Wiens\({^3}\),
Patrick D. Schloss \(^{1,\dagger}\)

\vspace{10mm}

\({^1}\) Department of Microbiology and Immunology, University of
Michigan, Ann Arbor, Michigan, USA

\({^2}\) Department of Computational Medicine and Bioinformatics,
University of Michigan, Ann Arbor, Michigan, USA

\({^3}\) Department of Electrical Engineering and Computer Science,
University of Michigan, Ann Arbor, Michigan, USA

\({^*}\) Current Affiliation:

\({^\#}\) Current Affiliation: Bristol Myers Squibb, Summit, New Jersey,
USA~

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

\vspace{10mm}

\textbf{observation format} (max 1200 words, 2 figures, 25 ref)

\newpage

\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Machine learning classification of disease based on the gut microbiome
often relies on clustering 16S rRNA gene sequences into operational
taxonomic units (OTUs) to quantify microbial composition. The standard
approach to clustering sequences into OTUs leverages the similarity of
the sequences to each other rather than to a reference database. The
abundance of each OTU is used to train a classification model. However,
OTU assignments depend on the sequences in the data set and therefore
can change if new data are added. This lack of stability complicates
classification because in order to use the model to classify additional
samples, the OTUs have to be reclustered to include the new sequences
and the model retrained with the new OTU clusters. A new reference-based
clustering algorithm, called OptiFit, addresses this issue by fitting
new sequences into existing OTUs. While OptiFit is proven to produce
high quality OTU clusters, it is unclear whether this method for fitting
new sequence data into existing OTUs will impact the performance of
classification models. We used OptiFit to cluster additional data into
existing OTU clusters and quantified model performance in classifying a
data set containing samples from patients with and without colonic
screen relevant neoplasias (SRN). We compared the performance of this
model to the standard procedure of clustering all the data together. We
found that both approaches performed equally well in classifying SRNs.
Moving forward, when OTUs are used in classification problems, OptiFit
can be used to avoid the need to retrain models using reclustered
sequences when classifying new samples.

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

There is great potential for using microbiome data to non-invasively
diagnose people. One of the challenges with using classification models
generated using the relative abundance of operational taxonomic unit
(OTU) data is that 16S rRNA gene sequences are assigned to OTUs based on
their similarity to other sequences in the dataset. If data are
generated from new patients seeking a diagnosis, then it would be
necessary to reassign sequences to OTUs and retrain the classification
model. Yet there is a desire to have a single, validated model that can
be deployed. To overcome this obstacle, we applied the OptiFit
clustering algorithm which fits new sequence data to existing OTUs
allowing the reuse of a consistent model. A random forest machine
learning model deployed using OptiFit performed as well as the
traditional reassignment and retraining approach. This result indicates
there is potential for developing machine learning classification models
based on OTU relative abundance data.

\newpage

Gut community composition is useful as a resource for machine learning
classification of diseases, including colorectal cancer
(\protect\hyperlink{ref-baxter2016}{1},
\protect\hyperlink{ref-zackular2014}{2}). Amplicon sequencing of the 16S
rRNA gene is a reliable tool for assessing the taxonomic composition of
microbial communities, which is the input to these models. Analysis of
16S rRNA gene sequence data generally relies on clustering of sequences
based on similarity into operational taxonomic units (OTUs). The process
of OTU clustering can either be reference-based or \emph{de novo}. The
quality of OTUs generated with reference-based clustering is generally
poor compared to those generated with \emph{de novo} clustering
(\protect\hyperlink{ref-westcott2015}{3}). While \emph{de novo}
clustering produces high-quality OTU clusters where sequences are
accurately grouped based on similarity thresholds, the resulting OTU
clusters depend on the data in the data set and the addition of new data
could change the overall OTU clusters. The unstable nature of OTU
clustering complicates deployment of machine learning models since
integration of additional data requires reclustering all the data and
retraining of the model. The ability to integrate new data into a
validated model without reclustering and retraining could allow for
deployment of a single model that new data can be continually added to.
Recently Sovacool \emph{et al} introduced OptiFit: a method for fitting
new sequence data into existing OTUs
(\protect\hyperlink{ref-sovacool2022}{4}). While OptiFit is proven to
effectively fit new sequence data to existing OTU clusters, it is
unknown if the use of OptiFit will have an impact on classification.
Here we tested the ability of OptiFit to cluster new sequence data into
existing OTU clusters for the purpose of classification of disease based
on gut microbiome composition.

We compared two approaches, one using all of the data to generate OTU
clusters and the other generating OTU clusters with a portion of the
data and then fitting the remaining sequence data to the existing OTUs
using OptiFit. In the first approach, all of the 16S rRNA sequence data
was \emph{de novo} clustered into OTUs with the OptiClust algorithm in
mothur (\protect\hyperlink{ref-westcott2017}{5}). The resulting
abundance data was then split into training and testing sets, where the
training set was used to tune hyperparameters and ultimately train the
model. The testing set was then classified with the model and the
performance of the model was quantified (Figure 1A). However, with this
methodology we would have to regenerate the OTU clusters and retrain the
model if we wanted to classify additional samples. The OptiFit algorithm
(\protect\hyperlink{ref-sovacool2022}{4}) addresses this problem by
enabling new sequences to be clustered into existing OTUs. The OptiFit
workflow is similar to the OptiClust workflow where the data was
clustered into OTUs and used to tune hyperparameters and ultimately
train the model. Then, we used OptiFit to fit sequence data of samples
not part of the original data set into the existing OTUs and used the
same model to classify the samples (Figure 1B). To test how the model
performance compares between these two methodologies, we used a publicly
available data set of 16S rRNA gene sequences from stool samples of
healthy subjects as well as subjects with SRN consisting of advanced
adenoma and carcinoma (\protect\hyperlink{ref-baxter2016}{1}). The data
set was randomly split into an 80\% train set and 20\% test set. For the
standard OptiClust workflow, the training and test sets were \emph{de
novo} clustered together into OTUs then the resulting abundance table
was split into the training and testing set. For the OptiFit workflow,
the train set was clustered \emph{de novo} into OTUs and the remaining
test set was fit to the OTU clusters using the OptiFit algorithm. For
both workflows, the abundance table of the train set was used to tune
hyperparameters and train a random forest model to classify SRN. The
test set was classified as either control or SRN using the trained
models.To account for variation depending on the split of the data, the
data set was randomly split 100 times and the process repeated for each
of the 100 data splits. By comparing the model performance of
classifying the samples in the test data set between the OptiFit and
OptiClust algorithms, we quantified the impact of using OptiFit on model
classification performance.

We first examined the quality of the resulting OTU clusters from the two
algorithms using the Matthews correlation coefficient (MCC). The MCC
score was quantified by examining all pairs of sequences and assessing
whether they belonged together in an OTU based on their similarity
(\protect\hyperlink{ref-westcott2017}{5}). MCC scores range between
negative one and one. A score of negative one means none of the
sequences in an OTU are within the similarity threshold and any
sequences within the similarity threshold are not in an OTU together. An
MCC score of zero essentially means the sequences are randomly
clustered. An MCC score of 1 means all sequences in an OTU are within
the similarity threshold and all sequence pairs within the similarity
threshold are in the same OTU. To ensure that OptiFit is appropriately
integrating new sequence data into the existing OTUs, we expected the
MCC scores produced by the OptiClust and OptiFit workflows to be
similar. Since the data was only clustered once in the OptiClust
workflow there was only one MCC score while the OptiFit workflow
produced an MCC score for the OTU clusters from each data split. Overall
the MCC scores were similar between OptiClust (MCC = 0.884) and OptiFit
(average MCC = 0.879). This indicated that OptiFit performed as well as
OptiClust when integrating new sequences into the existing OTUs.

After verifying that the quality of the OTUs was consistent between
OptiClust and OptiFit, we examined the model performance for classifying
samples in the held out test data set. To quantify model performance we
used the OTU relative abundances from the training data from the
OptiClust and OptiFit workflows to train a model to predict SRNs. Using
the predicted and actual diagnosis classification, we calculated the
area under the receiver operating characteristic curve (AUROC) for each
data split to quantify model performance. During cross-validation (CV)
training, the model performance was equivalent between the two
algorithms (p-value = 0.13, OptiClust mean CV AUROC = 0.694, OptiFit
mean CV AUROC = 0.697, Figure 2A). The trained model was then deployed
to classify the samples of the test data as control or SRN. The
performance on the test data was equivalent between the two algorithms
(p-value = 0.63, OptiClust mean test AUROC = 0.709, OptiFit mean test
AUROC = 0.712, Figure 2B,C) indicating that new data could be fit to
existing OTU clusters without impacting model performance.

We tested the ability of OptiFit to integrate new data into existing
OTUs for the purpose of machine learning classification using OTU
relative abundance. A potential problem with using OptiFit is that any
sequences in the new data that do not map to the existing OTU clusters
will be discarded resulting in a possible loss of information. However,
we demonstrated that OptiFit can be used to fit new sequence data into
existing OTU clusters and perform equally well in predicting SRN
compared to clustering all of the sequence data together. The ability to
integrate data from new samples into existing OTUs enables the
deployment of a single machine learning model. These results are based
on a single data set and disease. Further analysis is needed to
determine the number of samples that are necessary to build a robust
model capable of classifying diverse samples. A robust machine learning
model could be implemented as part of a non-invasive and low-cost aid in
diagnosing SRN and other diseases.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{\emph{Data Set.}} Raw 16S rRNA gene sequence data isolated from
human stool samples was downloaded from NCBI Sequence Read Archive
(accession no. SRP062005) (\protect\hyperlink{ref-baxter2016}{1},
\protect\hyperlink{ref-edgar2011}{6}). This data set contains stool
samples from a total of 490 subjects. For this analysis, samples from
subjects identified in the metadata as normal, high risk normal, or
adenoma were categorized as ``normal'' while samples from subjects
identified as advanced adenoma or carcinoma were categorized as ``screen
relevant neoplasia'' (SRN). The resulting data set consisted of 261
normal samples and 229 SRN samples.

\textbf{\emph{Data Processing.}} The full dataset was preprocessed with
mothur (v1.47) (\protect\hyperlink{ref-schloss2009}{7}) to join forward
and reverse reads, merge duplicate reads, align to the SILVA reference
database (v132) (\protect\hyperlink{ref-quast2013}{8}), precluster,
remove chimeras with UCHIME (\protect\hyperlink{ref-edgar2011}{6}),
assign taxonomy, and remove non-bacterial reads following the Schloss
Lab MiSeq standard operating procedure described on the mothur website
(\url{https://mothur.org/wiki/miseq_sop/}). 100 splits of the 490
samples were generated where 80\% of the samples (392 samples) were
randomly assigned to the training set and the remaining 20\% (98
samples) were assigned to the test set. Using 100 splits of the data
accounts for the variation that may be observed depending on the samples
that are in the training or test sets. Each sample was in the training
set an average of 80 times (SD=4.1) and the test set an average of 20
times (SD=4.1).

The data was processed through two workflows. First, the standard
workflow using the OptiClust algorithm
(\protect\hyperlink{ref-westcott2017}{5}). In this pathway, all of the
data was clustered together with OptiClust to generate OTUs and the
resulting abundance tables were split into the training and testing
sets. In the second workflow, the preprocessed data was split into the
training and testing sets. The training set was clustered into OTUs,
then the test set was fit to the OTUs of the training set using the
OptiFit algorithm (\protect\hyperlink{ref-sovacool2022}{4}). The OptiFit
algorithm was run with method open so that any sequences that did not
map to the existing OTU clusters would form new OTUs. For both pathways,
the shared files were sub-sampled to 10,000 reads per sample.

\textbf{\emph{Machine Learning.}} Machine learning using Random Forest
was conducted with the R package mikrompl (v 1.2.0)
(\protect\hyperlink{ref-topuxe7uoglu2021}{9}) to predict the diagnosis
(SRN or normal) for the samples in the test set for each data split. The
training set was preprocessed to normalize OTU counts (scale/center),
collapse correlated OTUs, and remove OTUs with zero-variance. The
preprocessing from the training set was then applied to the test set.
Any OTUs in the test set that were not in the training set were removed.
P values comparing model performance were calculated as previously
described (\protect\hyperlink{ref-topuxe7uoglu2020}{10}). The averaged
ROC curves were plotted by taking the average and standard deviation of
the sensitivity at each specificity value.

\textbf{\emph{Code Availability.}} The analysis workflow was implemented
in Snakemake (\protect\hyperlink{ref-koster2012}{11}) . Scripts for
analysis were written in R (\protect\hyperlink{ref-R2020}{12}) and GNU
bash (\protect\hyperlink{ref-GNUbash}{13}). The software used includes
mothur v1.47.0 (\protect\hyperlink{ref-schloss2009}{7}), RStudio
(\protect\hyperlink{ref-RStudio2019}{14}), the Tidyverse metapackage
(\protect\hyperlink{ref-wickham2019}{15}), R Markdown
(\protect\hyperlink{ref-xie_r_2018}{16}), the SRA toolkit
(\protect\hyperlink{ref-noauthor_sra-tools_nodate}{17}), and conda
(\protect\hyperlink{ref-noauthor_anaconda_2016}{18}). The complete
workflow and supporting files required to reproduce this study are
available at:
\url{https://github.com/SchlossLab/Armour_OptiFitGLNE_XXXX_2021}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This work was supported through a grant from the NIH (R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-baxter2016}{}}%
\CSLLeftMargin{1. }
\CSLRightInline{\textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers
MAM}, \textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}:37.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.}

\leavevmode\vadjust pre{\hypertarget{ref-zackular2014}{}}%
\CSLLeftMargin{2. }
\CSLRightInline{\textbf{Zackular JP}, \textbf{Rogers MAM},
\textbf{Ruffin MT,I}, \textbf{Schloss PD}. 2014. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research \textbf{7}:1112--1121.
doi:\href{https://doi.org/10.1158/1940-6207.CAPR-14-0129}{10.1158/1940-6207.CAPR-14-0129}.}

\leavevmode\vadjust pre{\hypertarget{ref-westcott2015}{}}%
\CSLLeftMargin{3. }
\CSLRightInline{\textbf{Westcott SL}, \textbf{Schloss PD}. 2015. De novo
clustering methods outperform reference-based methods for assigning 16S
rRNA gene sequences to operational taxonomic units. PeerJ
\textbf{3}:e1487.
doi:\href{https://doi.org/10.7717/peerj.1487}{10.7717/peerj.1487}.}

\leavevmode\vadjust pre{\hypertarget{ref-sovacool2022}{}}%
\CSLLeftMargin{4. }
\CSLRightInline{\textbf{Sovacool KL}, \textbf{Westcott SL},
\textbf{Mumphrey MB}, \textbf{Dotson GA}, \textbf{Schloss PD}. 2022.
OptiFit: An improved method for fitting amplicon sequences to existing
OTUs. mSphere \textbf{7}:e00916--21.
doi:\href{https://doi.org/10.1128/msphere.00916-21}{10.1128/msphere.00916-21}.}

\leavevmode\vadjust pre{\hypertarget{ref-westcott2017}{}}%
\CSLLeftMargin{5. }
\CSLRightInline{\textbf{Westcott SL}, \textbf{Schloss PD}. 2017.
OptiClust, an improved method for assigning amplicon-based sequence data
to operational taxonomic units. mSphere \textbf{2}:e00073--17.
doi:\href{https://doi.org/10.1128/mSphereDirect.00073-17}{10.1128/mSphereDirect.00073-17}.}

\leavevmode\vadjust pre{\hypertarget{ref-edgar2011}{}}%
\CSLLeftMargin{6. }
\CSLRightInline{\textbf{Edgar RC}, \textbf{Haas BJ}, \textbf{Clemente
JC}, \textbf{Quince C}, \textbf{Knight R}. 2011. UCHIME improves
sensitivity and speed of chimera detection. Bioinformatics
\textbf{27}:2194--2200.
doi:\href{https://doi.org/10.1093/bioinformatics/btr381}{10.1093/bioinformatics/btr381}.}

\leavevmode\vadjust pre{\hypertarget{ref-schloss2009}{}}%
\CSLLeftMargin{7. }
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Van Horn DJ}, \textbf{Weber CF}.
2009. Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.}

\leavevmode\vadjust pre{\hypertarget{ref-quast2013}{}}%
\CSLLeftMargin{8. }
\CSLRightInline{\textbf{Quast C}, \textbf{Pruesse E}, \textbf{Yilmaz P},
\textbf{Gerken J}, \textbf{Schweer T}, \textbf{Yarza P}, \textbf{Peplies
J}, \textbf{Glöckner FO}. 2013. The SILVA ribosomal RNA gene database
project: Improved data processing and web-based tools. Nucleic Acids
Research \textbf{41}:D590--D596.
doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.}

\leavevmode\vadjust pre{\hypertarget{ref-topuxe7uoglu2021}{}}%
\CSLLeftMargin{9. }
\CSLRightInline{\textbf{Topçuoğlu BD}, \textbf{Lapp Z}, \textbf{Sovacool
KL}, \textbf{Snitkin E}, \textbf{Wiens J}, \textbf{Schloss PD}. 2021.
mikropml: User-Friendly R Package for Supervised Machine Learning
Pipelines. Journal of Open Source Software \textbf{6}:3073.
doi:\href{https://doi.org/10.21105/joss.03073}{10.21105/joss.03073}.}

\leavevmode\vadjust pre{\hypertarget{ref-topuxe7uoglu2020}{}}%
\CSLLeftMargin{10. }
\CSLRightInline{\textbf{Topçuoğlu BD}, \textbf{Lesniak NA},
\textbf{Ruffin MT}, \textbf{Wiens J}, \textbf{Schloss PD}. 2020. A
framework for effective application of machine learning to
microbiome-based classification problems. mBio \textbf{11}:e00434--20.
doi:\href{https://doi.org/10.1128/mBio.00434-20}{10.1128/mBio.00434-20}.}

\leavevmode\vadjust pre{\hypertarget{ref-koster2012}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{\textbf{Koster J}, \textbf{Rahmann S}. 2012.
Snakemake--a scalable bioinformatics workflow engine. Bioinformatics
\textbf{28}:2520--2522.
doi:\href{https://doi.org/10.1093/bioinformatics/bts480}{10.1093/bioinformatics/bts480}.}

\leavevmode\vadjust pre{\hypertarget{ref-R2020}{}}%
\CSLLeftMargin{12. }
\CSLRightInline{\textbf{R Core Team}. 2020.
\href{https://www.R-project.org/}{R: A language and environment for
statistical computing}. R Foundation for Statistical Computing, Vienna,
Austria.}

\leavevmode\vadjust pre{\hypertarget{ref-GNUbash}{}}%
\CSLLeftMargin{13. }
\CSLRightInline{\textbf{GNU Project}.
\href{https://www.gnu.org/software/bash/\%20manual/bash.html/}{Bash
reference manual}.}

\leavevmode\vadjust pre{\hypertarget{ref-RStudio2019}{}}%
\CSLLeftMargin{14. }
\CSLRightInline{\textbf{RStudio Team}. 2019.
\href{http://www.rstudio.com/}{RStudio: Integrated development
environment for r}. RStudio, Inc., Boston, MA.}

\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
\CSLLeftMargin{15. }
\CSLRightInline{\textbf{Wickham H}, \textbf{Averick M}, \textbf{Bryan
J}, \textbf{Chang W}, \textbf{McGowan LD}, \textbf{François R},
\textbf{Grolemund G}, \textbf{Hayes A}, \textbf{Henry L}, \textbf{Hester
J}, \textbf{Kuhn M}, \textbf{Pedersen TL}, \textbf{Miller E},
\textbf{Bache SM}, \textbf{Müller K}, \textbf{Ooms J}, \textbf{Robinson
D}, \textbf{Seidel DP}, \textbf{Spinu V}, \textbf{Takahashi K},
\textbf{Vaughan D}, \textbf{Wilke C}, \textbf{Woo K}, \textbf{Yutani H}.
2019. Welcome to the Tidyverse. Journal of Open Source Software
\textbf{4}:1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.}

\leavevmode\vadjust pre{\hypertarget{ref-xie_r_2018}{}}%
\CSLLeftMargin{16. }
\CSLRightInline{\textbf{Xie Y}, \textbf{Allaire JJ}, \textbf{Grolemund
G}. 2018. R {Markdown}: {The Definitive Guide}. {Taylor \& Francis, CRC
Press}.}

\leavevmode\vadjust pre{\hypertarget{ref-noauthor_sra-tools_nodate}{}}%
\CSLLeftMargin{17. }
\CSLRightInline{{SRA}-{Tools} - {NCBI}.
http://ncbi.github.io/sra-tools/.}

\leavevmode\vadjust pre{\hypertarget{ref-noauthor_anaconda_2016}{}}%
\CSLLeftMargin{18. }
\CSLRightInline{2016. Anaconda {Software Distribution}. Anaconda
Documentation. Anaconda Inc.}

\end{CSLReferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\includegraphics{../exploratory/figures/figure1_a.pdf}

\includegraphics{../exploratory/figures/figure1_b.pdf}

\textbf{Figure 1: Workflows.} \textbf{A)} OptiClust workflow: The full
data set was clustered into OTUs using the OptiClust algorithm in
mothur. The data was then split into two sets where 80\% of the samples
were assigned to the training set and 20\% to the testing set. The
training set was preprocessed with mikropml to normalize values
(scale/center), collapse correlated features, and remove features with
zero-variance. Using mikropml, the training set was split into train and
validate sets to compare results using different hyperparameter
settings. The highest performing hyperparameter setting was then used to
train the model with the full training set. The preprocessing scale from
the training set was applied to the test data set, then the trained
model was used to classify the samples in the test set. Based on the
actual classification and predicted classification, the area under the
receiver operating characteristic curve (AUROC) was calculated to
summarize model performance. The entire process was repeated 100 times
to account for variability depending on the split of the data resulting
in a total of 100 AUROC values summarizing the performance of the
standard OptiClust workflow. \textbf{B)} OptiFit workflow: The data set
was first split into two sets where 80\% of the samples were assigned to
the training set and 20\% to the testing set. The training set was then
clustered into OTUs using the OptiClust algorithm in mothur. The
resulting abundance data was preprocessed with mikropml to normalize
values (scale/center), collapse correlated features, and remove features
with zero-variance. Using mikropml, the training set was split into
train and validate sets to compare results using different
hyperparameter settings. The highest performing hyperparameter setting
was then used to train the model with the full training set. The OptiFit
algorithm in mothur was used to cluster the left out testing data set
using the OTUs of the training set as a reference. The preprocessing
scale from the training set was applied to the test data set, then the
trained model was used to classify the samples in the test set. Based on
the actual classification and predicted classification, the area under
the receiver operating characteristic curve (AUROC) was calculated to
summarize model performance. The entire process was repeated 100 times
to account for variability depending on the split of the data resulting
in a total of 100 AUROC values summarizing the performance of the new
OptiFit workflow.

\newpage

\includegraphics{../exploratory/figures/figure2.pdf}

\textbf{Figure 2: Model Performance.} \textbf{A)} Area under the
receiver operating characteristic (AUROC) curve during cross-validation
for the OptiClust and OptiFit workflows. Mean and standard deviation of
the AUROC is represented by the black dot and whiskers. Mean AUROC is
printed to the right of the points. \textbf{B)} AUROC on the test data
for the OptiClust and OptiFit workflows. Mean and standard deviation of
the AUROC is represented by the black dot and whiskers. Mean AUROC is
printed to the right of the points. \textbf{C)} Receiver operating
characteristic (ROC)Averaged ROC curves

\end{document}