### 2023-06-15 submission/second_revision/manuscript.Rmd
second revision documents

---
title: "**Machine learning classification by fitting amplicon sequences to existing OTUs**"
output:
  rmarkdown::pdf_document:
    keep_tex: yes
geometry: margin=1.0in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \usepackage{multirow}
  - \usepackage[none]{hyphenat}
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
csl: mbio.csl
link-citations: true
---

```{r settings, eval=TRUE, echo=FALSE, cache=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("include" = FALSE)

format_decimal <- function(number,digits){
  return(round(number,digits=digits))
}

```

Running title: self-reference-based OTU clustering for ML classification

\vspace{10mm}

Courtney R. Armour${^1}$, Kelly L. Sovacool${^2}$, William L.
Close$^{1,*}$, Begüm D. Topçuoğlu$^{1,\#}$, Jenna Wiens${^3}$, Patrick
D. Schloss $^{1,\dagger}$

\vspace{10mm}

${^1}$ Department of Microbiology and Immunology, University of
Michigan, Ann Arbor, Michigan, USA

${^2}$ Department of Computational Medicine and Bioinformatics,
University of Michigan, Ann Arbor, Michigan, USA

${^3}$ Department of Electrical Engineering and Computer Science,
University of Michigan, Ann Arbor, Michigan, USA

${^*}$ Current Affiliation: Bio-Rad Laboratories, Hercules, California,
USA

${^\#}$ Current Affiliation: Bristol Myers Squibb, Summit, New Jersey,
USA

$\dagger$ To whom correspondence should be addressed:
[pschloss\@umich.edu](mailto:pschloss@umich.edu)

\newpage

\linenumbers

## Abstract

The ability to use 16S rRNA gene sequence data to train machine learning
classification models offers the opportunity diagnose patients based on
the composition of their microbiome. In some applications the taxonomic
resolution that provides the best models may require the use of *de
novo* OTUs whose composition changes when new data are added. We
previously developed a new reference-based approach, OptiFit, that fits
new sequence data to existing *de novo* OTUs without changing the
composition of the original OTUs. While OptiFit produces OTUs that are
as high quality as *de novo* OTUs, it is unclear whether this method for
fitting new sequence data into existing OTUs will impact the performance
of classification models relative to models trained and tested only
using *de novo* OTUs. We used OptiFit to cluster sequences into existing
OTUs and evaluated model performance in classifying a dataset containing
samples from patients with and without colonic screen relevant neoplasia
(SRN). We compared the performance of this model to standard methods
including *de novo* and database-reference-based clustering. We found
that using OptiFit performed as well or better in classifying SRNs.
OptiFit can streamline the process of classifying new samples by
avoiding the need to retrain models using reclustered sequences.

\newpage

## Importance

There is great potential for using microbiome data to aid in diagnosis.
A challenge with *de novo* OTU-based classification models is that 16S
rRNA gene sequences are often assigned to OTUs based on similarity to
other sequences in the dataset. If data are generated from new patients,
the old and new sequences must be reclustered to OTUs and the
classification model retrained. Yet there is a desire to have a single,
validated model that can be widely deployed. To overcome this obstacle,
we applied the OptiFit clustering algorithm to fit new sequence data to
existing OTUs allowing for reuse of the model. A random forest model
implemented using OptiFit performed as well as the traditional reassign
and retrain approach. This result shows that it is possible to train and
apply machine learning models based on OTU relative abundance data that
do not require retraining or the use of a reference database.

\newpage

There is increasing interest in training machine learning models to
diagnose diseases such as Crohn's disease and colorectal cancer using
the relative abundance of clusters of similar 16S rRNA gene sequences
[@baxter2016; @gevers2014]. These models have been used to identify
sequence clusters that are important for distinguishing between
individuals from different disease categories @duvallet2017. There is
also an opportunity to train models and apply them to classify samples
from new individuals. For example, a model for colorectal cancer could
be trained, "locked down", and applied to samples from new patients.

To apply these models to new samples the composition of the clusters
would need to be independent of the new data. For example, amplicon
sequence variants (ASVs) are defined without consideration of sequences
in other samples, phylotypes are defined by clustering sequences that
have the same taxonomy (e.g., to the same family) when classified using
a taxonomy database, and closed reference operational taxonomic units
(OTUs) are defined by mapping sequences to a collection of reference
OTUs. In contrast, *de novo* approaches cluster sequences based on their
similarity to other sequences in the dataset and can change when new
data are added. Although it would be preferrable to select an approach
that generates stable clusters, there may be cases where OTUs generated
by a *de novo* approach outperform those of the other taxonomic levels.
In fact, we recently trained machine learning models for classifying
patients with and without screen relevant neoplasias (SRNs) in their
colons and found that OTUs generated *de novo* using the OptiClust
algorithm performed better than those generated using ASVs or at higher
taxonomic levels @Armour2022.

It could be possible to construct reference OTUs and map new sequences
to those OTUs to attain similar performance as was seen with the
OptiClust-generated OTUs. The traditional approach to reference-based
clustering of sequences to OTUs has multiple drawbacks and
does not produce clusters as good as those generated using OptiClust
@sovacool2022. Sovacool *et al* recently described OptiFit, a method for
fitting new sequence data into existing OTUs that overcomes the
limitations of traditional reference-based clustering @sovacool2022.
OptiFit allows researchers to fit new data into existing OTUs defined
from the same dataset resulting in clusters that are as good as if they
had all been clustered with OptiClust. We tested whether
OptiClust-generated OTUs could be used to train models that were then
used to classify held out samples after clustering their sequences to
the model's OTUs using OptiFit.

To test how the model performance compared between using *de novo* and
reference-based clustering approaches, we used a publicly available
dataset of 16S rRNA gene sequences from stool samples of healthy
subjects (n = 226) as well as subjects with screen-relevant neoplasia
(SRN) consisting of advanced adenoma and carcinoma (n = 229)
@baxter2016. For the *de novo* workflows, the 16S rRNA sequence data
from all samples were clustered into OTUs using the OptiClust algorithm
in mothur @westcott2017 and the VSEARCH algorithm used in QIIME2
[@rognes2016; @bolyen2019]. For both algorithms, the resulting
abundance data was then split into training and testing sets, where the
training set was used to tune hyperparameters and ultimately train and
select the model. The model was applied to the testing set and
performance evaluated (Figure 1A). For traditional reference-based
clustering (database-reference-based), we used OptiFit to fit the
sequence data into OTUs based on the commonly used greengenes reference
database. To compare with another commonly used method, we also used
VSEARCH to map sequences to reference OTUs from the greengenes database
with the parameters used by QIIME2. Again, the data was then split into
training and testing sets, hyperparameters tuned, and performance
evaluated on the testing set (Figure 1B). In the OptiFit self-reference
workflow (self-reference-based), the data was split into a training and
a testing set. The training set was clustered into OTUs and used to
train a classification model. The OptiFit algorithm was used to fit
sequence data of samples not part of the training data into the training
OTUs and classified using the best hyperparameters (Figure 1C). For each
of the workflows the process was repeated for 100 random splits of the
data to account for variation caused by the choice of the random number
generator seed.

```{r mcc}
merged_mcc <- read_csv("../../results/ml/summary/merged_mcc.csv")

opticlust_denovo_mcc <- merged_mcc %>% 
  filter(algorithm == "opticlust_denovo") %>% 
  pull(mcc) %>% round(digits=3)

optifit_self_avg_mcc <- merged_mcc %>% 
  filter(algorithm == "optifit_self" & state=="combo") %>% 
  summarise(mean_mcc = mean(mcc),
            sd_mcc = sd(mcc)) %>% 
  round(digits=3)

optifit_gg_mcc <- merged_mcc %>% 
  filter(algorithm == "optifit_gg") %>% 
  pull(mcc) %>% round(digits=3)

vsearch_denovo_mcc <- merged_mcc %>% 
  filter(algorithm == "vsearch_denovo") %>% 
  pull(mcc) %>% round(digits=3)

vsearch_gg_mcc <- merged_mcc %>% 
  filter(algorithm == "vsearch_gg") %>% 
  pull(mcc) %>% round(digits=3)
```

```{r frac_mapped}
counts <- read_csv("../../results/tables/counts.csv")

avg_frac_mapped <- counts %>%
  mutate(frac_mapped = count/10000 * 100) %>%
  group_by(methods,set) %>%
  summarize(mean_frac = mean(frac_mapped),
            sd_frac = sd(frac_mapped),
            .groups="drop") %>% 
  filter(set == "test") 
```

We first examined the quality of the resulting OTU clusters from each
method using the Matthews correlation coefficient (MCC). MCC is an
objective metric used to measure OTU cluster quality based on the
similarity of all pairs of sequences and whether they are appropriately
clustered or not @westcott2015. We expected the MCC scores produced by
the OptiFit workflow to be similar to that of *de novo* clustering using
the OptiClust algorithm. In the OptiFit workflow the test data was fit
to the clustered training data for each of the 100 data splits resulting
in an MCC score for each split of the data. In the remaining workflows,
the data was only clustered once and then split into the training and
testing sets resulting in a single MCC score for each method. Indeed,
the MCC scores were similar between the OptiClust *de novo* (MCC =
`r opticlust_denovo_mcc`) and OptiFit self-reference workflows (average
MCC = `r optifit_self_avg_mcc %>% pull(mean_mcc)`, standard deviation =
`r optifit_self_avg_mcc %>% pull(sd_mcc)`). Consistent with prior
findings, the reference-based methods produced lower MCC scores (OptiFit
greengenes MCC = `r optifit_gg_mcc`; VSEARCH greengenes MCC =
`r vsearch_gg_mcc`) than the *de novo* methods (OptiClust *de novo* MCC
= `r opticlust_denovo_mcc`; VSEARCH *de novo* MCC =
`r vsearch_denovo_mcc`) @sovacool2022. Another metric we examined for
the OptiFit workflow was the fraction of sequences from the test set
that mapped to the reference OTUs. Since sequences that did not map to
reference OTUs were eliminated, if a high percentage of reads did not
map to an OTU we expected this loss of data to negatively impact
classification performance. We found that loss of data was not an issue
since on average
`r avg_frac_mapped %>% filter(methods == "optifit_self", set == "test") %>% pull(mean_frac) %>% round(digits=1)`%
(standard deviation =
`r avg_frac_mapped %>% filter(methods == "optifit_self", set == "test") %>% pull(sd_frac) %>% round(digits=1)`%)
of sequences in the subsampled test set mapped to the reference OTUs.
This number is higher than the average fraction of reads mapped in the
OptiFit greengenes workflow
(mean = `r avg_frac_mapped %>% filter(methods == "optifit_gg",set == "test") %>% pull(mean_frac) %>% round(digits=1)`%, standard deviation = 
`r avg_frac_mapped %>% filter(methods == "optifit_gg",set == "test") %>% pull(sd_frac) %>% round(digits=1)`).
These results indicate that the OptiFit self-reference method performed
as well as the OptiClust *de novo* method and better than using an
external database.

```{r performance}
performance <- read_csv("../../results/ml/summary/merged_performance.csv",
                    col_types = cols(algorithm=col_character(),
                                     method=col_character(),
                                     split=col_character(),
                                     .default=col_double()))

performance_summary <- performance %>% 
  group_by(algorithm) %>% 
  summarise(median_cv_AUC = round(median(cv_metric_AUC),digits=3),
            mean_cv_AUC = round(mean(cv_metric_AUC),digits=3),
            sd_cv_AUC = round(sd(cv_metric_AUC),digits=3),
            median_AUC = round(median(AUC),digits=3),
            mean_AUC = round(mean(AUC),digits=3),
            sd_AUC = round(sd(AUC),digits=3))

pvals <- read_csv("../../results/tables/pvalues.csv",col_types = cols(p_value = col_double(),
                                                                   .default = col_character()))

get_pval <- function(pvals,met,g1,g2){
  val <- pvals %>% 
    filter(metric == met) %>% 
    filter(group1 == g1 & group2 == g2 | group1 == g2 & group2 == g1) %>% 
    pull(p_value)

  format(round(val,digits = 3),nsmall=3)
}
```

We next assessed model performance using OTU relative abundances from
the training data from the workflows to train a model to predict SRNs
and used the model on the held-out data. Using the predicted and actual
diagnosis classification, we calculated the area under the receiver
operating characteristic curve (AUROC) for each data split. During
cross-validation (CV) training, the performance of the OptiFit
self-reference and OptiClust *de novo* models were not significantly
different (p-value =
`r get_pval(pvals, "cv_metric_AUC", "opticlust_denovo", "optifit_self")`;
Figure 2A), while performance for both VSEARCH methods was significantly
lower than the OptiClust *de novo*, OptiFit self, and OptiFit greengenes
methods (p-values \< 0.05). The trained model was then applied to the
test data classifying samples as either control or SRN. The VSEARCH
greengenes method performed slightly worse than the OptiClust *de novo*
method (p-value =
`r get_pval(pvals,"AUC","opticlust_denovo","vsearch_gg")`). However, the
performance on the test data for the OptiClust *de novo*, OptiFit
greengenes, OptiFit self-reference, and VSEARCH *de novo* approaches
were not significantly different (p-values \> 0.05; Figures 2B and 2C).
These results indicate that new data could be fit to existing OTU
clusters using OptiFit without impacting model performance.

Random forest machine learning models trained using OptiClust-generated
OTUs and tested using OptiFit-generated OTUs performed as well as a
model trained using entirely *de novo* OTU assignments. A potential
problem with reference-based clustering methods is that sequences that
do not map to the reference OTUs are discarded, resulting in a possible
loss of information. However, we demonstrated that the training samples
represented the most important OTUs for classifying samples. Missing
important OTUs is more of a risk when using a database-reference-based
method since not all environments are well represented in public
databases. Despite this and the lower quality OTUs, the
database-reference-based approach performed as well as the models
generated using OptiFit. This likely indicates that the sequences that
were important to the model were well characterized by the greengenes
reference OTUs. However, a less well studied system may not be as well
characterized by a reference-database which would make the ability to
utilize one's own data a reference an exciting possibility. Our results
highlight that OptiFit overcomes a significant limitation with machine
learning models trained using *de novo* OTUs. This is an important
result for those applications where models trained using *de novo* OTUs
outperform models generated using methods that produce clusters that do
not depend on which sequences are included in the dataset.

## Materials and Methods

**Dataset.** Raw 16S rRNA gene sequence data from the V4 region were
previously generated from human stool samples. Sequences were downloaded
from the NCBI Sequence Read Archive (accession no. SRP062005)
@baxter2016. This dataset contains stool samples from 490 subjects. For
this analysis, samples from subjects identified in the metadata as
normal, high risk normal, or adenoma were categorized as "normal", while
samples from subjects identified as advanced adenoma or carcinoma were
categorized as "screen relevant neoplasia" (SRN). The resulting dataset
consisted of 261 normal samples and 229 SRN samples.

**Data processing.** The full dataset was preprocessed with mothur
(v1.47) @schloss2009 to join forward and reverse reads, merge duplicate
reads, align to the SILVA reference database (v132) @quast2013,
precluster, remove chimeras with UCHIME @edgar2011, assign taxonomy, and
remove non-bacterial reads following the Schloss Lab MiSeq standard
operating procedure described on the mothur website
(<https://mothur.org/wiki/miseq_sop/>). 100 splits of the 490 samples
were generated where 80% of the samples (392 samples) were randomly
assigned to the training set and the remaining 20% (98 samples) were
assigned to the test set. Using 100 splits of the data accounts for the
variation that may be observed depending on the samples that are in the
training or test sets. Each sample was in the training set an average of
80 times (standard deviation = 4.1) and the test set an average of 20
times (standard deviation = 4.1).

***Reference-based workflows.***

1.  OptiFit Self: The preprocess data was split into the training and
    testing sets. The training set was clustered into OTUs using
    OptiClust, then the test set was fit to the OTUs of the training set
    using the OptiFit algorithm @sovacool2022. The OptiFit algorithm was
    run with method open so that any sequences that did not map to the
    existing OTU clusters would form new OTUs. The data was then
    subsampled to 10,000 reads and any novel OTUs from the test set were
    removed. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
2.  OptiFit Greengenes: Reference sequences from the Greengenes database
    v13_8\_99 @desantis2006 were downloaded and processed with mothur by
    trimming to the V4 region and clustered *de novo* with OptiClust
    @westcott2017. The preprocessed data was fit to the clustered
    reference data using OptiFit with the method open to allow any
    sequences that did not map to the existing reference clusters would
    form new OTUs. The data was then subsampled to 10,000 reads and any
    novel OTUs from the test set were removed. The dataset was then
    split into two sets where 80% of the samples were assigned to the
    training set and 20% to the testing set. This process was repeated
    for each of the 100 splits resulting in 100 training and testing
    datasets.
3.  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
    v2.15.2 @rognes2016 directly to unprocessed Greengenes 97% OTU
    reference alignment consistent with how VSEARCH is typically used by
    the QIIME2 software for reference-based clustering @bolyen2019. The
    data was then subsampled to 10,000 reads and any novel OTUs from the
    test set were removed. The dataset was then split into two sets
    where 80% of the samples were assigned to the training set and 20%
    to the testing set. This process was repeated for each of the 100
    splits resulting in 100 training and testing datasets.

***De novo workflows.***

4.  OptiClust *de novo*: All the preprocessed data was clustered
    together with OptiClust @westcott2017 to generate OTUs. The data was
    subsampled to 10,000 reads per sample and the resulting abundance
    tables were split into the training and testing sets. The process
    was repeated for each of the 100 splits resulting in 100 training
    and testing datasets.
5.  VSEARCH *de novo*: All the preprocessed data was clustered using
    VSEARCH v2.15.2 @rognes2016 with 97% identity and then subsampled to
    10,000 reads per sample. The process was repeated for each of the
    100 splits resulting in 100 training and testing datasets for both
    workflows.

**Machine Learning.** A random forest model was trained with the R
package mikrompl (v 1.2.0) @topçuoglu2021 to predict the diagnosis (SRN
or normal) for the samples in the test set for each data split. The
training set was preprocessed to normalize OTU counts (scale and
center), collapse correlated OTUs, and remove OTUs with zero variance.
The preprocessing from the training set was then applied to the test
set. Any OTUs in the test set that were not in the training set were
removed. P-values comparing model performance were calculated as
previously described @topçuoglu2020. The averaged ROC curves were
plotted by taking the average and standard deviation of the sensitivity
at each specificity value.

**Code Availability.** The analysis workflow was implemented in
Snakemake @koster2012. Scripts for analysis were written in R @R2020 and
GNU bash @GNUbash. The software used includes mothur v1.47.0
@schloss2009, VSEARCH v2.15.2 @rognes2016, RStudio @RStudio2019, the
Tidyverse metapackage @wickham2019, R Markdown @xie_r_2018, the SRA
toolkit @noauthor_sra-tools_nodate, and conda @noauthor_anaconda_2016.
The complete workflow and supporting files required to reproduce this
study are available at:
<https://github.com/SchlossLab/Armour_OptiFitGLNE_XXXX_2023>

## Acknowledgments

This work was supported through a grant from the NIH (R01CA215574).

\newpage

## References

```{=tex}
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
```
\noindent

::: {#refs}
:::

```{=tex}
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
```
\newpage

## Figure Legends

**Figure1: Overview of clustering workflows.** The *de novo* and
database-reference-based workflows were conducted using two approaches:
OptiClust with mothur and VSEARCH as is used in the QIIME pipeline.

**Figure 2: Model performance of OptiFit self-reference workflow is as
good or better than other methods.** **A)** Area under the receiver
operating characteristic (AUROC) curve during cross-validation (train)
for the various workflows. **B)** AUROC on the test data for the various
workflows. The mean and standard deviation of the AUROC is represented
by the black dot and whiskers in panels A and B. The mean AUROC is
printed below the points. **C)** Averaged receiver operating
characteristic (ROC) curves. Lines represent the average true positive
rate for the range of false positive rates.