### 2023-06-15 submission/second_revision/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\usepackage{multirow}
\usepackage[none]{hyphenat}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Machine learning classification by fitting amplicon sequences to existing OTUs},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Machine learning classification by fitting amplicon
sequences to existing OTUs}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

Running title: self-reference-based OTU clustering for ML classification

\vspace{10mm}

Courtney R. Armour\({^1}\), Kelly L. Sovacool\({^2}\), William L.
Close\(^{1,*}\), Begüm D. Topçuoğlu\(^{1,\#}\), Jenna Wiens\({^3}\),
Patrick D. Schloss \(^{1,\dagger}\)

\vspace{10mm}

\({^1}\) Department of Microbiology and Immunology, University of
Michigan, Ann Arbor, Michigan, USA

\({^2}\) Department of Computational Medicine and Bioinformatics,
University of Michigan, Ann Arbor, Michigan, USA

\({^3}\) Department of Electrical Engineering and Computer Science,
University of Michigan, Ann Arbor, Michigan, USA

\({^*}\) Current Affiliation: Bio-Rad Laboratories, Hercules,
California, USA

\({^\#}\) Current Affiliation: Bristol Myers Squibb, Summit, New Jersey,
USA

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

\newpage

\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The ability to use 16S rRNA gene sequence data to train machine learning
classification models offers the opportunity diagnose patients based on
the composition of their microbiome. In some applications the taxonomic
resolution that provides the best models may require the use of \emph{de
novo} OTUs whose composition changes when new data are added. We
previously developed a new reference-based approach, OptiFit, that fits
new sequence data to existing \emph{de novo} OTUs without changing the
composition of the original OTUs. While OptiFit produces OTUs that are
as high quality as \emph{de novo} OTUs, it is unclear whether this
method for fitting new sequence data into existing OTUs will impact the
performance of classification models relative to models trained and
tested only using \emph{de novo} OTUs. We used OptiFit to cluster
sequences into existing OTUs and evaluated model performance in
classifying a dataset containing samples from patients with and without
colonic screen relevant neoplasia (SRN). We compared the performance of
this model to standard methods including \emph{de novo} and
database-reference-based clustering. We found that using OptiFit
performed as well or better in classifying SRNs. OptiFit can streamline
the process of classifying new samples by avoiding the need to retrain
models using reclustered sequences.

\newpage

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

There is great potential for using microbiome data to aid in diagnosis.
A challenge with \emph{de novo} OTU-based classification models is that
16S rRNA gene sequences are often assigned to OTUs based on similarity
to other sequences in the dataset. If data are generated from new
patients, the old and new sequences must be reclustered to OTUs and the
classification model retrained. Yet there is a desire to have a single,
validated model that can be widely deployed. To overcome this obstacle,
we applied the OptiFit clustering algorithm to fit new sequence data to
existing OTUs allowing for reuse of the model. A random forest model
implemented using OptiFit performed as well as the traditional reassign
and retrain approach. This result shows that it is possible to train and
apply machine learning models based on OTU relative abundance data that
do not require retraining or the use of a reference database.

\newpage

There is increasing interest in training machine learning models to
diagnose diseases such as Crohn's disease and colorectal cancer using
the relative abundance of clusters of similar 16S rRNA gene sequences
(\protect\hyperlink{ref-baxter2016}{1},
\protect\hyperlink{ref-gevers2014}{2}). These models have been used to
identify sequence clusters that are important for distinguishing between
individuals from different disease categories
(\protect\hyperlink{ref-duvallet2017}{3}). There is also an opportunity
to train models and apply them to classify samples from new individuals.
For example, a model for colorectal cancer could be trained, ``locked
down'', and applied to samples from new patients.

To apply these models to new samples the composition of the clusters
would need to be independent of the new data. For example, amplicon
sequence variants (ASVs) are defined without consideration of sequences
in other samples, phylotypes are defined by clustering sequences that
have the same taxonomy (e.g., to the same family) when classified using
a taxonomy database, and closed reference operational taxonomic units
(OTUs) are defined by mapping sequences to a collection of reference
OTUs. In contrast, \emph{de novo} approaches cluster sequences based on
their similarity to other sequences in the dataset and can change when
new data are added. Although it would be preferrable to select an
approach that generates stable clusters, there may be cases where OTUs
generated by a \emph{de novo} approach outperform those of the other
taxonomic levels. In fact, we recently trained machine learning models
for classifying patients with and without screen relevant neoplasias
(SRNs) in their colons and found that OTUs generated \emph{de novo}
using the OptiClust algorithm performed better than those generated
using ASVs or at higher taxonomic levels
(\protect\hyperlink{ref-Armour2022}{4}).

It could be possible to construct reference OTUs and map new sequences
to those OTUs to attain similar performance as was seen with the
OptiClust-generated OTUs. The traditional approach to reference-based
clustering of sequences to OTUs has multiple drawbacks and does not
produce clusters as good as those generated using OptiClust
(\protect\hyperlink{ref-sovacool2022}{5}). Sovacool \emph{et al}
recently described OptiFit, a method for fitting new sequence data into
existing OTUs that overcomes the limitations of traditional
reference-based clustering (\protect\hyperlink{ref-sovacool2022}{5}).
OptiFit allows researchers to fit new data into existing OTUs defined
from the same dataset resulting in clusters that are as good as if they
had all been clustered with OptiClust. We tested whether
OptiClust-generated OTUs could be used to train models that were then
used to classify held out samples after clustering their sequences to
the model's OTUs using OptiFit.

To test how the model performance compared between using \emph{de novo}
and reference-based clustering approaches, we used a publicly available
dataset of 16S rRNA gene sequences from stool samples of healthy
subjects (n = 226) as well as subjects with screen-relevant neoplasia
(SRN) consisting of advanced adenoma and carcinoma (n = 229)
(\protect\hyperlink{ref-baxter2016}{1}). For the \emph{de novo}
workflows, the 16S rRNA sequence data from all samples were clustered
into OTUs using the OptiClust algorithm in mothur 1 and the VSEARCH
algorithm used in QIIME2 (\protect\hyperlink{ref-rognes2016}{6},
\protect\hyperlink{ref-bolyen2019}{7}). For both algorithms, the
resulting abundance data was then split into training and testing sets,
where the training set was used to tune hyperparameters and ultimately
train and select the model. The model was applied to the testing set and
performance evaluated (Figure 1A). For traditional reference-based
clustering (database-reference-based), we used OptiFit to fit the
sequence data into OTUs based on the commonly used greengenes reference
database. To compare with another commonly used method, we also used
VSEARCH to map sequences to reference OTUs from the greengenes database
with the parameters used by QIIME2. Again, the data was then split into
training and testing sets, hyperparameters tuned, and performance
evaluated on the testing set (Figure 1B). In the OptiFit self-reference
workflow (self-reference-based), the data was split into a training and
a testing set. The training set was clustered into OTUs and used to
train a classification model. The OptiFit algorithm was used to fit
sequence data of samples not part of the training data into the training
OTUs and classified using the best hyperparameters (Figure 1C). For each
of the workflows the process was repeated for 100 random splits of the
data to account for variation caused by the choice of the random number
generator seed.

We first examined the quality of the resulting OTU clusters from each
method using the Matthews correlation coefficient (MCC). MCC is an
objective metric used to measure OTU cluster quality based on the
similarity of all pairs of sequences and whether they are appropriately
clustered or not (\protect\hyperlink{ref-westcott2015}{8}). We expected
the MCC scores produced by the OptiFit workflow to be similar to that of
\emph{de novo} clustering using the OptiClust algorithm. In the OptiFit
workflow the test data was fit to the clustered training data for each
of the 100 data splits resulting in an MCC score for each split of the
data. In the remaining workflows, the data was only clustered once and
then split into the training and testing sets resulting in a single MCC
score for each method. Indeed, the MCC scores were similar between the
OptiClust \emph{de novo} (MCC = 0.884) and OptiFit self-reference
workflows (average MCC = 0.879, standard deviation = 0.002). Consistent
with prior findings, the reference-based methods produced lower MCC
scores (OptiFit greengenes MCC = 0.786; VSEARCH greengenes MCC = 0.531)
than the \emph{de novo} methods (OptiClust \emph{de novo} MCC = 0.884;
VSEARCH \emph{de novo} MCC = 0.641)
(\protect\hyperlink{ref-sovacool2022}{5}). Another metric we examined
for the OptiFit workflow was the fraction of sequences from the test set
that mapped to the reference OTUs. Since sequences that did not map to
reference OTUs were eliminated, if a high percentage of reads did not
map to an OTU we expected this loss of data to negatively impact
classification performance. We found that loss of data was not an issue
since on average 99.8\% (standard deviation = 0.7\%) of sequences in the
subsampled test set mapped to the reference OTUs. This number is higher
than the average fraction of reads mapped in the OptiFit greengenes
workflow (mean = 96.8\%, standard deviation = 3.5). These results
indicate that the OptiFit self-reference method performed as well as the
OptiClust \emph{de novo} method and better than using an external
database.

We next assessed model performance using OTU relative abundances from
the training data from the workflows to train a model to predict SRNs
and used the model on the held-out data. Using the predicted and actual
diagnosis classification, we calculated the area under the receiver
operating characteristic curve (AUROC) for each data split. During
cross-validation (CV) training, the performance of the OptiFit
self-reference and OptiClust \emph{de novo} models were not
significantly different (p-value = 0.066; Figure 2A), while performance
for both VSEARCH methods was significantly lower than the OptiClust
\emph{de novo}, OptiFit self, and OptiFit greengenes methods (p-values
\textless{} 0.05). The trained model was then applied to the test data
classifying samples as either control or SRN. The VSEARCH greengenes
method performed slightly worse than the OptiClust \emph{de novo} method
(p-value = 0.030). However, the performance on the test data for the
OptiClust \emph{de novo}, OptiFit greengenes, OptiFit self-reference,
and VSEARCH \emph{de novo} approaches were not significantly different
(p-values \textgreater{} 0.05; Figures 2B and 2C). These results
indicate that new data could be fit to existing OTU clusters using
OptiFit without impacting model performance.

Random forest machine learning models trained using OptiClust-generated
OTUs and tested using OptiFit-generated OTUs performed as well as a
model trained using entirely \emph{de novo} OTU assignments. A potential
problem with reference-based clustering methods is that sequences that
do not map to the reference OTUs are discarded, resulting in a possible
loss of information. However, we demonstrated that the training samples
represented the most important OTUs for classifying samples. Missing
important OTUs is more of a risk when using a database-reference-based
method since not all environments are well represented in public
databases. Despite this and the lower quality OTUs, the
database-reference-based approach performed as well as the models
generated using OptiFit. This likely indicates that the sequences that
were important to the model were well characterized by the greengenes
reference OTUs. However, a less well studied system may not be as well
characterized by a reference-database which would make the ability to
utilize one's own data a reference an exciting possibility. Our results
highlight that OptiFit overcomes a significant limitation with machine
learning models trained using \emph{de novo} OTUs. This is an important
result for those applications where models trained using \emph{de novo}
OTUs outperform models generated using methods that produce clusters
that do not depend on which sequences are included in the dataset.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\textbf{Dataset.} Raw 16S rRNA gene sequence data from the V4 region
were previously generated from human stool samples. Sequences were
downloaded from the NCBI Sequence Read Archive (accession no. SRP062005)
(\protect\hyperlink{ref-baxter2016}{1}). This dataset contains stool
samples from 490 subjects. For this analysis, samples from subjects
identified in the metadata as normal, high risk normal, or adenoma were
categorized as ``normal'', while samples from subjects identified as
advanced adenoma or carcinoma were categorized as ``screen relevant
neoplasia'' (SRN). The resulting dataset consisted of 261 normal samples
and 229 SRN samples.

\textbf{Data processing.} The full dataset was preprocessed with mothur
(v1.47) (\protect\hyperlink{ref-schloss2009}{9}) to join forward and
reverse reads, merge duplicate reads, align to the SILVA reference
database (v132) (\protect\hyperlink{ref-quast2013}{10}), precluster,
remove chimeras with UCHIME (\protect\hyperlink{ref-edgar2011}{11}),
assign taxonomy, and remove non-bacterial reads following the Schloss
Lab MiSeq standard operating procedure described on the mothur website
(\url{https://mothur.org/wiki/miseq_sop/}). 100 splits of the 490
samples were generated where 80\% of the samples (392 samples) were
randomly assigned to the training set and the remaining 20\% (98
samples) were assigned to the test set. Using 100 splits of the data
accounts for the variation that may be observed depending on the samples
that are in the training or test sets. Each sample was in the training
set an average of 80 times (standard deviation = 4.1) and the test set
an average of 20 times (standard deviation = 4.1).

\textbf{\emph{Reference-based workflows.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  OptiFit Self: The preprocess data was split into the training and
  testing sets. The training set was clustered into OTUs using
  OptiClust, then the test set was fit to the OTUs of the training set
  using the OptiFit algorithm (\protect\hyperlink{ref-sovacool2022}{5}).
  The OptiFit algorithm was run with method open so that any sequences
  that did not map to the existing OTU clusters would form new OTUs. The
  data was then subsampled to 10,000 reads and any novel OTUs from the
  test set were removed. This process was repeated for each of the 100
  splits resulting in 100 training and testing datasets.
\item
  OptiFit Greengenes: Reference sequences from the Greengenes database
  v13\_8\_99 (\protect\hyperlink{ref-desantis2006}{12}) were downloaded
  and processed with mothur by trimming to the V4 region and clustered
  \emph{de novo} with OptiClust

  \begin{enumerate}
  \def\labelenumii{\arabic{enumii}.}
  \tightlist
  \item
    The preprocessed data was fit to the clustered reference data using
    OptiFit with the method open to allow any sequences that did not map
    to the existing reference clusters would form new OTUs. The data was
    then subsampled to 10,000 reads and any novel OTUs from the test set
    were removed. The dataset was then split into two sets where 80\% of
    the samples were assigned to the training set and 20\% to the
    testing set. This process was repeated for each of the 100 splits
    resulting in 100 training and testing datasets.
  \end{enumerate}
\item
  VSEARCH Greengenes: Preprocessed data was clustered using VSEARCH
  v2.15.2 (\protect\hyperlink{ref-rognes2016}{6}) directly to
  unprocessed Greengenes 97\% OTU reference alignment consistent with
  how VSEARCH is typically used by the QIIME2 software for
  reference-based clustering (\protect\hyperlink{ref-bolyen2019}{7}).
  The data was then subsampled to 10,000 reads and any novel OTUs from
  the test set were removed. The dataset was then split into two sets
  where 80\% of the samples were assigned to the training set and 20\%
  to the testing set. This process was repeated for each of the 100
  splits resulting in 100 training and testing datasets.
\end{enumerate}

\textbf{\emph{De novo workflows.}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  OptiClust \emph{de novo}: All the preprocessed data was clustered
  together with OptiClust 1 to generate OTUs. The data was subsampled to
  10,000 reads per sample and the resulting abundance tables were split
  into the training and testing sets. The process was repeated for each
  of the 100 splits resulting in 100 training and testing datasets.
\item
  VSEARCH \emph{de novo}: All the preprocessed data was clustered using
  VSEARCH v2.15.2 (\protect\hyperlink{ref-rognes2016}{6}) with 97\%
  identity and then subsampled to 10,000 reads per sample. The process
  was repeated for each of the 100 splits resulting in 100 training and
  testing datasets for both workflows.
\end{enumerate}

\textbf{Machine Learning.} A random forest model was trained with the R
package mikrompl (v 1.2.0)
(\protect\hyperlink{ref-topuxe7uoglu2021}{13}) to predict the diagnosis
(SRN or normal) for the samples in the test set for each data split. The
training set was preprocessed to normalize OTU counts (scale and
center), collapse correlated OTUs, and remove OTUs with zero variance.
The preprocessing from the training set was then applied to the test
set. Any OTUs in the test set that were not in the training set were
removed. P-values comparing model performance were calculated as
previously described (\protect\hyperlink{ref-topuxe7uoglu2020}{14}). The
averaged ROC curves were plotted by taking the average and standard
deviation of the sensitivity at each specificity value.

\textbf{Code Availability.} The analysis workflow was implemented in
Snakemake (\protect\hyperlink{ref-koster2012}{15}). Scripts for analysis
were written in R (\protect\hyperlink{ref-R2020}{16}) and GNU bash
(\protect\hyperlink{ref-GNUbash}{17}). The software used includes mothur
v1.47.0 (\protect\hyperlink{ref-schloss2009}{9}), VSEARCH v2.15.2
(\protect\hyperlink{ref-rognes2016}{6}), RStudio
(\protect\hyperlink{ref-RStudio2019}{18}), the Tidyverse metapackage
(\protect\hyperlink{ref-wickham2019}{19}), R Markdown
(\protect\hyperlink{ref-xie_r_2018}{20}), the SRA toolkit
(\protect\hyperlink{ref-noauthor_sra-tools_nodate}{21}), and conda
(\protect\hyperlink{ref-noauthor_anaconda_2016}{22}). The complete
workflow and supporting files required to reproduce this study are
available at:
\url{https://github.com/SchlossLab/Armour_OptiFitGLNE_XXXX_2023}

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

This work was supported through a grant from the NIH (R01CA215574).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-baxter2016}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers
MAM}, \textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}:37.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.}

\leavevmode\vadjust pre{\hypertarget{ref-gevers2014}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Gevers D}, \textbf{Kugathasan S}, \textbf{Denson
LA}, \textbf{Vázquez-Baeza Y}, \textbf{Van~Treuren W}, \textbf{Ren B},
\textbf{Schwager E}, \textbf{Knights D}, \textbf{Song SJ},
\textbf{Yassour M}, \textbf{Morgan XC}, \textbf{Kostic AD}, \textbf{Luo
C}, \textbf{González A}, \textbf{McDonald D}, \textbf{Haberman Y},
\textbf{Walters T}, \textbf{Baker S}, \textbf{Rosh J}, \textbf{Stephens
M}, \textbf{Heyman M}, \textbf{Markowitz J}, \textbf{Baldassano R},
\textbf{Griffiths A}, \textbf{Sylvester F}, \textbf{Mack D}, \textbf{Kim
S}, \textbf{Crandall W}, \textbf{Hyams J}, \textbf{Huttenhower C},
\textbf{Knight R}, \textbf{Xavier RJ}. 2014. The treatment-naive
microbiome in new-onset crohn{'}s disease. Cell Host \& Microbe
\textbf{15}:382--392.
doi:\href{https://doi.org/10.1016/j.chom.2014.02.005}{10.1016/j.chom.2014.02.005}.}

\leavevmode\vadjust pre{\hypertarget{ref-duvallet2017}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{\textbf{Duvallet C}, \textbf{Gibbons SM}, \textbf{Gurry
T}, \textbf{Irizarry RA}, \textbf{Alm EJ}. 2017. Meta-analysis of gut
microbiome studies identifies disease-specific and shared responses.
Nature Communications \textbf{8}:1784.
doi:\href{https://doi.org/10.1038/s41467-017-01973-8}{10.1038/s41467-017-01973-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-Armour2022}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{\textbf{Armour CR}, \textbf{Topçuoğlu BD},
\textbf{Garretto A}, \textbf{Schloss PD}. 2022. A goldilocks principle
for the gut microbiome: Taxonomic resolution matters for
microbiome-based classification of colorectal cancer. MBio
\textbf{13}:e0316121.}

\leavevmode\vadjust pre{\hypertarget{ref-sovacool2022}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{\textbf{Sovacool KL}, \textbf{Westcott SL},
\textbf{Mumphrey MB}, \textbf{Dotson GA}, \textbf{Schloss PD}. 2022.
OptiFit: An improved method for fitting amplicon sequences to existing
OTUs. mSphere \textbf{7}:e00916--21.
doi:\href{https://doi.org/10.1128/msphere.00916-21}{10.1128/msphere.00916-21}.}

\leavevmode\vadjust pre{\hypertarget{ref-rognes2016}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{\textbf{Rognes T}, \textbf{Flouri T}, \textbf{Nichols
B}, \textbf{Quince C}, \textbf{Mahé F}. 2016. VSEARCH: a versatile open
source tool for metagenomics. PeerJ \textbf{4}:e2584.
doi:\href{https://doi.org/10.7717/peerj.2584}{10.7717/peerj.2584}.}

\leavevmode\vadjust pre{\hypertarget{ref-bolyen2019}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{\textbf{Bolyen E}, \textbf{Rideout JR}, \textbf{Dillon
MR}, \textbf{Bokulich NA}, \textbf{Abnet CC}, \textbf{Al-Ghalith GA},
\textbf{Alexander H}, \textbf{Alm EJ}, \textbf{Arumugam M},
\textbf{Asnicar F}, \textbf{Bai Y}, \textbf{Bisanz JE},
\textbf{Bittinger K}, \textbf{Brejnrod A}, \textbf{Brislawn CJ},
\textbf{Brown CT}, \textbf{Callahan BJ}, \textbf{Caraballo-Rodríguez
AM}, \textbf{Chase J}, \textbf{Cope EK}, \textbf{Da Silva R},
\textbf{Diener C}, \textbf{Dorrestein PC}, \textbf{Douglas GM},
\textbf{Durall DM}, \textbf{Duvallet C}, \textbf{Edwardson CF},
\textbf{Ernst M}, \textbf{Estaki M}, \textbf{Fouquier J},
\textbf{Gauglitz JM}, \textbf{Gibbons SM}, \textbf{Gibson DL},
\textbf{Gonzalez A}, \textbf{Gorlick K}, \textbf{Guo J},
\textbf{Hillmann B}, \textbf{Holmes S}, \textbf{Holste H},
\textbf{Huttenhower C}, \textbf{Huttley GA}, \textbf{Janssen S},
\textbf{Jarmusch AK}, \textbf{Jiang L}, \textbf{Kaehler BD},
\textbf{Kang KB}, \textbf{Keefe CR}, \textbf{Keim P}, \textbf{Kelley
ST}, \textbf{Knights D}, \textbf{Koester I}, \textbf{Kosciolek T},
\textbf{Kreps J}, \textbf{Langille MGI}, \textbf{Lee J}, \textbf{Ley R},
\textbf{Liu Y-X}, \textbf{Loftfield E}, \textbf{Lozupone C},
\textbf{Maher M}, \textbf{Marotz C}, \textbf{Martin BD},
\textbf{McDonald D}, \textbf{McIver LJ}, \textbf{Melnik AV},
\textbf{Metcalf JL}, \textbf{Morgan SC}, \textbf{Morton JT},
\textbf{Naimey AT}, \textbf{Navas-Molina JA}, \textbf{Nothias LF},
\textbf{Orchanian SB}, \textbf{Pearson T}, \textbf{Peoples SL},
\textbf{Petras D}, \textbf{Preuss ML}, \textbf{Pruesse E},
\textbf{Rasmussen LB}, \textbf{Rivers A}, \textbf{Robeson MS},
\textbf{Rosenthal P}, \textbf{Segata N}, \textbf{Shaffer M},
\textbf{Shiffer A}, \textbf{Sinha R}, \textbf{Song SJ}, \textbf{Spear
JR}, \textbf{Swafford AD}, \textbf{Thompson LR}, \textbf{Torres PJ},
\textbf{Trinh P}, \textbf{Tripathi A}, \textbf{Turnbaugh PJ},
\textbf{Ul-Hasan S}, \textbf{Hooft JJJ van der}, \textbf{Vargas F},
\textbf{Vázquez-Baeza Y}, \textbf{Vogtmann E}, \textbf{Hippel M von},
\textbf{Walters W}, \textbf{Wan Y}, \textbf{Wang M}, \textbf{Warren J},
\textbf{Weber KC}, \textbf{Williamson CHD}, \textbf{Willis AD},
\textbf{Xu ZZ}, \textbf{Zaneveld JR}, \textbf{Zhang Y}, \textbf{Zhu Q},
\textbf{Knight R}, \textbf{Caporaso JG}. 2019. Reproducible,
interactive, scalable and extensible microbiome data science using QIIME
2. Nature Biotechnology \textbf{37}:852--857.
doi:\href{https://doi.org/10.1038/s41587-019-0209-9}{10.1038/s41587-019-0209-9}.}

\leavevmode\vadjust pre{\hypertarget{ref-westcott2015}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{\textbf{Westcott SL}, \textbf{Schloss PD}. 2015. De novo
clustering methods outperform reference-based methods for assigning 16S
rRNA gene sequences to operational taxonomic units. PeerJ
\textbf{3}:e1487.
doi:\href{https://doi.org/10.7717/peerj.1487}{10.7717/peerj.1487}.}

\leavevmode\vadjust pre{\hypertarget{ref-schloss2009}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Van Horn DJ}, \textbf{Weber CF}.
2009. Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.}

\leavevmode\vadjust pre{\hypertarget{ref-quast2013}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{\textbf{Quast C}, \textbf{Pruesse E}, \textbf{Yilmaz P},
\textbf{Gerken J}, \textbf{Schweer T}, \textbf{Yarza P}, \textbf{Peplies
J}, \textbf{Glöckner FO}. 2013. The SILVA ribosomal RNA gene database
project: Improved data processing and web-based tools. Nucleic Acids
Research \textbf{41}:D590--D596.
doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.}

\leavevmode\vadjust pre{\hypertarget{ref-edgar2011}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{\textbf{Edgar RC}, \textbf{Haas BJ}, \textbf{Clemente
JC}, \textbf{Quince C}, \textbf{Knight R}. 2011. UCHIME improves
sensitivity and speed of chimera detection. Bioinformatics
\textbf{27}:2194--2200.
doi:\href{https://doi.org/10.1093/bioinformatics/btr381}{10.1093/bioinformatics/btr381}.}

\leavevmode\vadjust pre{\hypertarget{ref-desantis2006}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{\textbf{DeSantis TZ}, \textbf{Hugenholtz P},
\textbf{Larsen N}, \textbf{Rojas M}, \textbf{Brodie EL}, \textbf{Keller
K}, \textbf{Huber T}, \textbf{Dalevi D}, \textbf{Hu P}, \textbf{Andersen
GL}. 2006. Greengenes, a chimera-checked 16S rRNA gene database and
workbench compatible with ARB. Applied and Environmental Microbiology
\textbf{72}:5069--5072.
doi:\href{https://doi.org/10.1128/AEM.03006-05}{10.1128/AEM.03006-05}.}

\leavevmode\vadjust pre{\hypertarget{ref-topuxe7uoglu2021}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{\textbf{Topçuoğlu BD}, \textbf{Lapp Z}, \textbf{Sovacool
KL}, \textbf{Snitkin E}, \textbf{Wiens J}, \textbf{Schloss PD}. 2021.
mikropml: User-Friendly R Package for Supervised Machine Learning
Pipelines. Journal of Open Source Software \textbf{6}:3073.
doi:\href{https://doi.org/10.21105/joss.03073}{10.21105/joss.03073}.}

\leavevmode\vadjust pre{\hypertarget{ref-topuxe7uoglu2020}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{\textbf{Topçuoğlu BD}, \textbf{Lesniak NA},
\textbf{Ruffin MT}, \textbf{Wiens J}, \textbf{Schloss PD}. 2020. A
framework for effective application of machine learning to
microbiome-based classification problems. mBio \textbf{11}:e00434--20.
doi:\href{https://doi.org/10.1128/mBio.00434-20}{10.1128/mBio.00434-20}.}

\leavevmode\vadjust pre{\hypertarget{ref-koster2012}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{\textbf{Koster J}, \textbf{Rahmann S}. 2012.
Snakemake--a scalable bioinformatics workflow engine. Bioinformatics
\textbf{28}:2520--2522.
doi:\href{https://doi.org/10.1093/bioinformatics/bts480}{10.1093/bioinformatics/bts480}.}

\leavevmode\vadjust pre{\hypertarget{ref-R2020}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{\textbf{R Core Team}. 2020.
\href{https://www.R-project.org/}{R: A language and environment for
statistical computing}. R Foundation for Statistical Computing, Vienna,
Austria.}

\leavevmode\vadjust pre{\hypertarget{ref-GNUbash}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{\textbf{GNU Project}.
\href{https://www.gnu.org/software/bash/\%20manual/bash.html/}{Bash
reference manual}.}

\leavevmode\vadjust pre{\hypertarget{ref-RStudio2019}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{\textbf{RStudio Team}. 2019.
\href{http://www.rstudio.com/}{RStudio: Integrated development
environment for r}. RStudio, Inc., Boston, MA.}

\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{\textbf{Wickham H}, \textbf{Averick M}, \textbf{Bryan
J}, \textbf{Chang W}, \textbf{McGowan LD}, \textbf{François R},
\textbf{Grolemund G}, \textbf{Hayes A}, \textbf{Henry L}, \textbf{Hester
J}, \textbf{Kuhn M}, \textbf{Pedersen TL}, \textbf{Miller E},
\textbf{Bache SM}, \textbf{Müller K}, \textbf{Ooms J}, \textbf{Robinson
D}, \textbf{Seidel DP}, \textbf{Spinu V}, \textbf{Takahashi K},
\textbf{Vaughan D}, \textbf{Wilke C}, \textbf{Woo K}, \textbf{Yutani H}.
2019. Welcome to the Tidyverse. Journal of Open Source Software
\textbf{4}:1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.}

\leavevmode\vadjust pre{\hypertarget{ref-xie_r_2018}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{\textbf{Xie Y}, \textbf{Allaire JJ}, \textbf{Grolemund
G}. 2018. R {Markdown}: {The Definitive Guide}. {Taylor \& Francis, CRC
Press}.}

\leavevmode\vadjust pre{\hypertarget{ref-noauthor_sra-tools_nodate}{}}%
\CSLLeftMargin{21. }%
\CSLRightInline{{SRA}-{Tools} - {NCBI}.
http://ncbi.github.io/sra-tools/.}

\leavevmode\vadjust pre{\hypertarget{ref-noauthor_anaconda_2016}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{2016. Anaconda {Software Distribution}. Anaconda
Documentation. Anaconda Inc.}

\end{CSLReferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage

\hypertarget{figure-legends}{%
\subsection{Figure Legends}\label{figure-legends}}

\textbf{Figure1: Overview of clustering workflows.} The \emph{de novo}
and database-reference-based workflows were conducted using two
approaches: OptiClust with mothur and VSEARCH as is used in the QIIME
pipeline.

\textbf{Figure 2: Model performance of OptiFit self-reference workflow
is as good or better than other methods.} \textbf{A)} Area under the
receiver operating characteristic (AUROC) curve during cross-validation
(train) for the various workflows. \textbf{B)} AUROC on the test data
for the various workflows. The mean and standard deviation of the AUROC
is represented by the black dot and whiskers in panels A and B. The mean
AUROC is printed below the points. \textbf{C)} Averaged receiver
operating characteristic (ROC) curves. Lines represent the average true
positive rate for the range of false positive rates.

\end{document}